---
title: "Meeting a journal's figure requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meeting a journal's figure requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(ggplot2)
library(figspec)
```

## The problem

A journal's figure requirements are specific and they differ between
publishers. Cell Press asks for figures at 85, 114 or 174 mm and type between
6 and 8 pt. Frontiers asks for 85 or 180 mm and type no smaller than 8 pt.
PLOS ONE asks for a width between 66.8 and 190.5 mm and type between 8 and
12 pt. Nothing in a normal plotting workflow makes any of that visible.

## Checking before you save

```{r}
p <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(title = "Fuel economy", x = "Weight (1000 lbs)", y = "Miles per gallon")

check_journal(p, "plos_one")
```

The type size check is the one that catches people. ggplot2's default base size
is 11 pt, which puts the title at 13.2 pt and axis text at 8.8 pt. PLOS ONE
caps figure type at 12 pt, so the title fails.

## Fixing it

`theme_journal()` enforces the journal's typography without imposing a look of
its own. It takes a base theme and pins every text element inside the
journal's stated range.

```{r}
p_fixed <- p + theme_journal("plos_one")
check_journal(p_fixed, "plos_one")
```

## Saving at the right size

This is the part that matters most, and it is easy to get wrong.

Type size in a figure is absolute. If you save a figure at 180 mm and the
journal places it in an 85 mm column, the whole figure is scaled to 47% and
your compliant 8 pt label arrives at 3.8 pt. Saving at the journal's own
column width is what stops that happening.

```{r, eval = FALSE}
ggsave_journal("figure_1.tiff", p_fixed, journal = "plos_one", column = "single")
```

`ggsave_journal()` reads the width from the registry, defaults the resolution
to the journal's stated minimum, warns if the format is not one the journal
accepts, and checks the file it just wrote.

## Reading the registry

```{r}
journals(discipline = "physics")
```

```{r}
journal_spec("cell_press")
```

## What figspec will not tell you

figspec reports four outcomes, and only one of them is a problem you must fix.

`unspecified` means the publisher does not state that requirement. That is not
a pass. If Frontiers says nothing about maximum file size, figspec will not
reassure you that your file size is fine.

`unknown` means the requirement exists but the input cannot answer it. Type
size cannot be recovered from a TIFF, so checking a saved raster file reports
type size as unknown. Check the plot object before you save it, where the
answer is knowable.

```{r}
r <- check_journal(p_fixed, "frontiers")
r[, c("check", "status")]
```

## Provenance

Every entry names its source and the date it was read.

```{r}
spec <- journal_spec("frontiers")
spec$source_url
spec$verified_on
```

Author guidelines change. Treat the publisher's current page as the authority,
and open an issue if you find an entry that has drifted.
