---
title: "Getting started with figspec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with figspec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(ggplot2)
library(figspec)
```

## Two jobs

figspec builds a figure to a specification, exports it at exactly that size and
resolution, then checks the result and says where it falls short.

A specification can come from two places, and the package does two jobs
accordingly.

**A journal's published requirements.** Name a target journal and the width,
resolution, file format, typography and colour rules come from its author
guidelines, as data, with the page they were read from and the date. This is
what the rest of this article covers.

**A size you choose yourself.** No journal need be involved. The most useful
case is setting the *plot panel* rather than the image, so a set of figures
shares one plot area whatever their axis labels do — something R cannot
otherwise express. That has its own article: `vignette("panels")`.

You can also write a specification by hand, or register a house style, and
check against it exactly as though it were a journal. Wherever this article
says "journal", a specification of your own works the same way.

## The problem journals create

Journals publish precise rules for the figures you submit, and the rules
differ. Cell Press asks for 85, 114 or 174 mm wide with type between 6 and 8 pt.
Science asks for 57, 121 or 184 mm with type no smaller than 5 pt. PNAS caps
type at 12 pt and asks that numerical axes reach zero. Nature wants panel
labels upright; the Royal Society wants them italicised.

Nothing in an ordinary plotting workflow makes any of that visible, and most
people find out at the production stage, after acceptance.

## Bring the journal in as you plot

The simplest way to meet a journal's requirements is to build the figure to
them from the start. Add the journal the way you would add a colour scale:

```{r}
fitted <- ggplot(mtcars, aes(wt, mpg, colour = factor(cyl), shape = factor(cyl))) +
  geom_point() +
  labs(title = "Fuel economy", x = "Weight (1000 lbs)", y = "Miles per gallon") +
  fit_journal("cell_press")

report <- fig_check(fitted, "cell_press", column = "single")
knitr::kable(report[, c("check", "actual", "status")], row.names = FALSE)
```

That one line carries the journal's typography, its stated line weights, its
structural rules, and colours and shapes chosen to survive what the journal
does to a figure in production.

## Checking a figure you have already drawn

Type size lives in the plot object. Once a figure becomes a TIFF it is pixels,
and the point sizes have gone with it, so the moment to check is while you
still hold the plot.

```{r}
p <- ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_point() +
  labs(title = "Fuel economy", x = "Weight (1000 lbs)", y = "Miles per gallon",
       colour = "Cylinders")

fig_check(p, "cell_press", column = "single")
```

Two real failures in an ordinary ggplot, from defaults nobody thinks about.
ggplot2's base size is 11 pt, which puts text between 8.8 and 13.2 pt against
Cell Press's 6 to 8. And ggplot2's default palette contains a red and a green,
which Cell Press states should not be used together.

## Reading the report

Each requirement gets one of four outcomes, and only one of them is a problem
you must fix.

| Outcome | Meaning |
|---|---|
| <span class="fs-status fs-pass">pass</span> | Meets the requirement |
| <span class="fs-status fs-fail">fail</span> | Breaches it |
| <span class="fs-status fs-unspecified">unspecified</span> | The publisher does not state this. Nothing can be concluded |
| <span class="fs-status fs-unknown">unknown</span> | The requirement exists, but this input cannot answer it |

`unspecified` splits further, and the distinction matters:

```{r}
r <- fig_check(p, "frontiers")
knitr::kable(
  r[r$check %in% c("File size", "Height"), c("check", "requirement", "status")],
  row.names = FALSE
)
```

Frontiers' file-size rule was read and confirmed absent, so it is *not
specified by publisher*. Its height rule has not been resolved into a number by
anyone, so it is *not yet harvested for this journal*. The first is a fact
about Frontiers. The second is a fact about this registry, and saying so is the
difference between a registry you can trust and one that is confidently wrong.

## Fixing it

`theme_journal()` applies every requirement a theme can express: the font
family, the type floor and ceiling, stated line weights, and structural rules
such as Nature's requirement that axis lines and tick marks be drawn.

```{r}
fixed <- p +
  scale_colour_figspec("cividis") +
  scale_shape_manual(values = figspec_shapes(3)) +
  aes(shape = factor(cyl)) +
  theme_journal("cell_press")

r <- fig_check(fixed, "cell_press", column = "single")
knitr::kable(
  r[r$status %in% c("pass", "fail"), c("check", "actual", "status")],
  row.names = FALSE
)
```

## Colour, and why shape matters

Three colour findings are reported, and which of them can *fail* depends on
what the publisher actually states.

```{r}
knitr::kable(
  check_colour_safety(p, "cell_press")[, c("check", "actual", "status")],
  row.names = FALSE
)
```

Red-and-green is a **requirement** for Cell Press because Cell Press states it.
Greyscale separability is a requirement for the Royal Society, which prints in
black and white by default. Colour-vision safety is reported for every journal
and stays `unspecified`, since it is good practice that publishers leave to
your judgement.

Note the last row. If colours merge but shapes differ, the figure still reads:

```{r}
knitr::kable(
  check_colour_safety(fixed, "royal_society")[, c("check", "actual", "status")],
  row.names = FALSE
)
```

`figspec_palettes()` lists the palettes that ship with the package, each with
its source. Cividis is safe both under colour vision deficiency and in
greyscale; Okabe-Ito is safe under colour vision deficiency but **not** in
greyscale, because two of its colours sit at almost the same lightness.

## Saving at the right size

Sizes here are canvas sizes: the image file is the width the journal states.
That is what compliance means, and it is what `fig_save()` does by default.
There is a second width — the plot area inside the image — which the journal
says nothing about and which is what makes a set of figures look even or
uneven. `vignette("panels")` covers it.


This is the part that is easiest to get wrong. Type size is absolute. Save a
figure at 180 mm, have the journal place it in an 85 mm column, and everything
is scaled to 47%: a compliant 8 pt label arrives at 3.8 pt.

```{r, eval = FALSE}
fig_save("figure_1.tiff", fixed, journal = "cell_press", column = "single")
```

`fig_save()` takes the width from the registry, defaults the resolution
to the journal's stated minimum, picks a device that can render the required
font, and re-checks the file it wrote.

Column names come from the journal itself:

```{r}
fig_columns("science")
fig_columns("cell_press")
```

## Which resolution rule applies

Publishers set different resolutions for different kinds of artwork, and the
gap is large: 300 dpi generally, against 600 to 1200 dpi for line art. Twelve
entries in the registry state a separate line-art figure.

Line art means **monochrome**, and the distinction carries real weight.
Springer defines it as a
"Black and white graphic with no shading"; ACS and IEEE both write "black and
white line art"; Taylor & Francis and OUP both write "monochrome". In prepress
it is a one-bit image, and it carries the highest resolution bar precisely
because sharp one-bit edges alias badly when sampled too coarsely.

So a colour figure is *not* line art, and asking 1200 dpi of it would be the
wrong advice. `suggest_art_type()` classifies what the plot actually contains:

```{r}
line_art <- ggplot(mtcars, aes(wt, mpg)) + geom_point(colour = "black")
suggest_art_type(line_art, "bmj")
```

ggplot2's default bar fill is `#595959`, a mid grey, so a default bar chart is
grayscale art:

```{r}
suggest_art_type(ggplot(mtcars, aes(factor(cyl))) + geom_bar())
```

`fig_check()` checks against the general minimum unless you say otherwise,
and mentions the line-art rule only when a plot really is pure black and white
and the journal holds line art to more. It does not change the verdict for you.

## Checking a whole submission

```{r, eval = FALSE}
check_submission("figures/", "cell_press")
#> ✔ Figure_1.tiff  single  all requirements met
#> ✖ Figure_2.tiff  double  failed: Resolution, File format
```

Files are checked as files, so geometry, format, resolution and file size can
be judged but type size cannot. That is why step one happens before saving.

## In R Markdown or Quarto

A large share of academic figures never pass through `ggsave()` at all: they
come out of a knitr chunk at whatever `fig.width` happens to be.

```{r}
figspec_chunk_opts("cell_press", "double")
```

Put that in a setup chunk with `figspec_knitr_setup("cell_press", "double")`
and every figure in the document is emitted at the journal's size.

## When the paper is rejected

The next journal has different rules, and they can be incompatible: Cell Press
wants type between 6 and 8 pt, PLOS ONE between 8 and 12.

```{r, eval = FALSE}
refit_journal(my_plots, journal = "plos_one", outdir = "figures_plos/")
```

`refit_journal()` re-themes and re-exports the set. It works from plot objects
and refuses saved files, because type size cannot be recovered from a finished
TIFF and a re-fitted file could not be trusted.

## Other display items

Journals set rules for other display items too.

```{r}
graphical_abstract_spec("rsc")
```

`table_spec()` and `media_spec()` do the same for tables and for supplementary
video and audio.

## Where figspec hands the judgement back to you

Where a publisher never said what compliance means, figspec reports the
silence. It leaves a pixel size in pixels until a resolution arrives to convert
it with. It reports a recommendation as a recommendation. And when type size is
the question it asks for the plot object, since a saved raster has already
turned that into pixels.

Each of those is a moment where the honest answer is "check this yourself", and
figspec says so plainly instead of filling the gap with a guess.

The same rule governs what figspec will *not* claim. Panel consistency across a
set of figures is reported, because it is measurable and it is usually what an
author is trying to fix — but never as a pass or a failure, because no
publisher asks for it. A green tick against an invented rule would be the same
error as inventing the rule.

## Where to go next

* `vignette("panels")` — sizing by the plot panel rather than the image, and
  making a set of figures match.
* `vignette("journals")` — every entry in the registry, and how complete each
  one is.
* `vignette("options")` — every function and what its arguments do.
