---
title: "Getting started with figspec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with figspec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(ggplot2)
library(figspec)
```

## Choose the requirements for your figure or table

A clear and accurate figure or table may still need changes before it can be
submitted, published or placed in a report. A figure may be too wide, have too
few pixels for print or contain text that will be unreadable at its final size.
A table may need to remain editable, use a particular page orientation or
follow rules for type and borders. Figures saved at the same width can also
look uneven when their plotting areas do not line up.

Start by providing figspec with the requirements or specifications for the
finished work. You can do this in three ways:

* **Use an existing publication profile.** Select your intended journal or
  publisher from figspec's registry. figspec loads the recorded requirements
  and shows where they came from and when they were last verified.
* **Write and use your own requirements.** Create a specification for a
  report, presentation, thesis, organisation or production workflow in a named
  R list. You can use it immediately or save it for reuse.
* **Set the required dimensions directly.** Give the exact image or
  plotting-panel size when size is the only constraint.

Whichever route you take, the workflow is the same: define the requirements →
build the figure or table to meet them → export it correctly → verify the
finished file → repeat when the work or its requirements change.

### Using your own requirements

For your own project, write the requirements in a named R list. Here, a
quarterly report accepts PNG or PDF figures that are 160 mm wide, at least
300 dpi, with text no smaller than 9 pt and lines no thinner than 0.5 pt:

```{r}
report_spec <- list(
  name = "Quarterly outcomes report",
  columns = list(full = 160),
  formats = c("png", "pdf"),
  dpi_min = 300,
  font_min_pt = 9,
  min_line_pt = 0.5
)
```

Use `report_spec` wherever a function asks for a specification,
including `fig_apply_spec()`, `theme_spec()`, `fig_check()` and `fig_save()`.
Those functions accept either a registry name or your own list, so no journal
is required.

To use the same requirements in a later R session, save them in a project YAML
file and load that file when the project starts:

```{r save-project-specification}
project_spec_file <- file.path(tempdir(), "project-specifications.yml")

spec_save(
  report_spec,
  project_spec_file,
  id = "quarterly_outcomes_report"
)
spec_load(project_spec_file)

saved_report_spec <- spec_get("quarterly_outcomes_report")
fig_width(saved_report_spec, "full")
```

In a real project, use a path such as `"project-specifications.yml"` and keep
the file with the analysis. `spec_save()` creates or updates the file safely;
`spec_load()` makes its entries available in the current session. Users do not
need to change figspec or add files to the installed package.

A **specification** lists the requirements the finished work must meet. A
**house style** sets visual choices, such as a reusable ggplot2 theme. They can
be used together, but they are kept separate so a preferred appearance is not
mistaken for a requirement. Use `style_register()` to name that appearance for
the current session and `style_save()` to keep it for later sessions.

### Reusing your own requirements and visual styles

The two kinds of reusable addition serve different purposes:

| If you want to reuse... | It should contain... | Save and restore it with... |
|---|---|---|
| A specification | Requirements that a finished figure or table must meet | `spec_save()` and `spec_load()` |
| A house style | Preferred ggplot2 appearance, such as the theme, grid and legend position | `style_save()` and `style_load()` |

Specifications are saved as readable YAML. House styles are saved as RDS
because they contain R theme objects. For example, the report specification
above has already been saved and loaded. A visual style can be preserved in a
separate file:

```{r save-project-style}
style_register(
  "quarterly_report",
  theme_minimal() +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom"
    ),
  description = "Quarterly report figures"
)

project_style_file <- file.path(tempdir(), "project-figure-styles.rds")
style_save(project_style_file)
style_remove("quarterly_report")
style_load(project_style_file)
style_list()[style_list()$name == "quarterly_report", ]
```

Once both files have been loaded, their names can be combined whenever a plot
is built:

```{r apply-saved-spec-and-style}
ggplot(ggplot2::mpg, aes(displ, hwy, colour = factor(cyl))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(
    x = "Engine displacement (litres)",
    y = "Highway miles per gallon",
    colour = "Cylinders"
  ) +
  fig_apply_spec(
    "quarterly_outcomes_report",
    style = "quarterly_report"
  )
```

For a project or research team, a simple layout is:

```text
project/
├── config/
│   └── figspec/
│       ├── specifications.yml
│       └── styles.rds
└── analysis.R
```

Load the files at the beginning of an analysis, in a project setup script or
in the first code chunk of an R Markdown or Quarto document:

```{r project-setup-example, eval = FALSE}
library(figspec)

spec_load(file.path("config", "figspec", "specifications.yml"))
style_load(file.path("config", "figspec", "styles.rds"))
```

These paths are relative to the project directory, so they work on every team
member's computer even when the project is stored in a different location.

Each team member has a separate R installation and a separate local copy of the
project. Git records changes to the project files, while GitHub or another Git
hosting service stores the shared repository. The team shares the figspec setup
in the same way that it shares analysis code:

1. Add `.Rprofile`, `config/figspec/specifications.yml` and
   `config/figspec/styles.rds` to the project repository.
2. Commit the files and push the commit to GitHub.
3. Each team member clones the repository to their computer the first time they
   join the project.
4. When someone updates a specification or style, they commit and push the
   change. Other team members pull it into their own project copy.
5. When a team member opens the project, their own R installation reads the
   local `.Rprofile` and loads the local copies of the specification and style
   files.
6. If a team member pulls an update while R is already running, they restart R
   or run `spec_load()` and `style_load()` again.

The team is not sharing one running copy of R, and `.Rprofile` is not sent
directly from one computer to another. GitHub holds the shared project files;
Git places a copy on each team member's computer. Every computer must have
figspec installed. A project that uses `renv` can also record the package
version so that the team uses the same release.

An explicit setup script is usually easiest to see and debug. If every R
session in the project should load the files automatically, put the following
in the project's `.Rprofile`:

```{r rprofile-example, eval = FALSE}
if (requireNamespace("figspec", quietly = TRUE)) {
  specification_file <- file.path(
    "config", "figspec", "specifications.yml"
  )
  style_file <- file.path("config", "figspec", "styles.rds")

  if (file.exists(specification_file)) {
    figspec::spec_load(specification_file)
  }
  if (file.exists(style_file)) {
    figspec::style_load(style_file)
  }
}
```

Because `.Rprofile` contains R code that runs when the project starts, review
changes to it before accepting them and never store passwords, access tokens or
other secrets in it. Load an RDS style file only from a repository controlled
by people you trust. Stored theme functions are blocked by default and require
the explicit `allow_functions = TRUE` option.

An organisation with many specifications, styles and helper functions can put
them in a small internal R package that imports figspec. That is the most
maintainable way to provide a larger shared extension across many projects.

For exact panel sizing without any specification, see the
[panel-sizing guide](panels.html). It demonstrates why equal image widths do
not guarantee equal plotting areas and how to make a set align.

## Using a journal or publisher profile

Journals and publishers make a useful example because they give exact
measurements and do not all use the same ones. Cell Press asks for widths of
85, 114 or 174 mm and type between 6 and 8 pt. Science uses widths of 57, 121
or 184 mm and type no smaller than 5 pt. PNAS caps type at 12 pt, while Nature
and the Royal Society use different panel-label styles.

Each publisher is designing a different page, so the differences are
reasonable. The difficulty is that a normal ggplot gives no warning when a
figure breaks one of these rules. The problem may only be discovered after
submission or during production.

## Apply the requirements while you build

It is easier to make changes while the plot is still editable. Add a
publication profile much like a ggplot2 scale:

```{r}
fitted <- ggplot(ggplot2::mpg,
                 aes(displ, hwy, colour = factor(cyl), shape = factor(cyl))) +
  geom_point(size = 2.1, alpha = 0.8, stroke = 0.6) +
  labs(title = "Engine size and highway fuel economy",
       x = "Engine displacement (litres)", y = "Highway miles per gallon",
       colour = "Cylinders", shape = "Cylinders") +
  fig_apply_spec("cell_press")

report <- fig_check(fitted, "cell_press", column = "single")
knitr::kable(report[, c("check", "actual", "status")], row.names = FALSE)
```

`fig_apply_spec()` applies the typography and structural rules that a ggplot2
theme can control. It also uses accessible colours and, when the plot maps a
variable to shape, shapes designed to remain distinguishable in print.

### What figspec changes

The two figures below use the same data, variables and labels. Both are drawn
on a Cell Press single-column canvas, which is 85 mm wide. The website displays
them at a larger size so that the details are easy to compare; the underlying
figure dimensions and measurements remain 85 mm. ggplot2's default text sizes
do not adjust to the requirements of that smaller canvas, so labels that look
fine on screen may be too large, too small or too crowded in the final figure.

```{r before, fig.width = 85/25.4, fig.height = 62/25.4, dpi = 200, out.width = "75%", fig.alt = "A default ggplot at 85 mm wide, with type too large and a red-green palette."}
plain <- ggplot(ggplot2::mpg, aes(displ, hwy, colour = factor(cyl))) +
  geom_point() +
  labs(title = "Engine size and highway fuel economy",
       x = "Engine displacement (litres)", y = "Highway miles per gallon",
       colour = "Cylinders")

plain
```

ggplot2's default base size is 11 pt, which gives this figure text ranging from
8.8 to 13.2 pt. Cell Press allows 6 to 8 pt. The default palette also includes
red and green together, which Cell Press does not allow, and colour is the only
way to tell the four series apart.

```{r after, fig.width = 85/25.4, fig.height = 62/25.4, dpi = 200, out.width = "75%", fig.alt = "The same figure with fig_apply_spec applied: smaller type, Okabe-Ito colours, and a different shape per series."}
plain +
  aes(shape = factor(cyl)) +
  # Give the shape the same label as the colour, so ggplot2 merges the two
  # into one legend rather than drawing a second.
  labs(shape = "Cylinders") +
  fig_apply_spec("cell_press")
```

The text now falls within the journal's range. The Okabe-Ito palette avoids a
red-green pairing and remains distinguishable for people with common forms of
colour-vision deficiency. Each series also has its own shape, so the figure
still works in black and white. The data have not changed.

## Checking a figure you have already drawn

Check the plot before saving whenever possible. At that stage, R still knows
the text sizes, colours, layers and theme settings. In a finished TIFF, text is
stored as pixels and its original point size cannot be recovered reliably.

```{r}
p <- plain

fig_check(p, "cell_press", column = "single")
```

The report finds two failures caused by ordinary ggplot2 defaults. The text is
larger than Cell Press permits, and the default palette uses red and green
together.

## Reading the report

Each requirement receives one of five results. A `fail` identifies something
that must be fixed. An `invalid` result means the file could not be assessed
reliably.

| Outcome | Meaning |
|---|---|
| <span class="fs-status fs-pass">pass</span> | Meets the requirement |
| <span class="fs-status fs-fail">fail</span> | Does not meet the requirement |
| <span class="fs-status fs-unspecified">unspecified</span> | No requirement is stated, so there is nothing to test |
| <span class="fs-status fs-unknown">unknown</span> | A requirement exists, but the plot or file does not contain enough information to test it |
| <span class="fs-status fs-fail">invalid</span> | The file is damaged or its contents do not match its extension |

When no value is available for a requirement, figspec distinguishes between
two reasons:

```{r}
r <- fig_check(p, "frontiers")
knitr::kable(
  r[r$check %in% c("File size", "Height"), c("check", "requirement", "status")],
  row.names = FALSE
)
```

Frontiers does not publish a file-size limit in the guidance that was reviewed,
so figspec reports *not specified by publisher*. A maximum figure height has
not yet been entered in the registry, so it reports *not yet harvested for this
journal*. The first describes the publisher's guidance; the second describes
the current state of the registry. Keeping them separate prevents missing data
from being mistaken for confirmation that no rule exists.

## Fixing what failed

`theme_spec()` applies the requirements that ggplot2's theme system can
control, including the font family, minimum and maximum text sizes, theme line
widths, and structural rules such as Nature's requirement to show axis lines
and tick marks.

```{r}
fixed <- p +
  scale_colour_figspec("cividis") +
  scale_shape_manual(values = figspec_shapes(4)) +
  aes(shape = factor(cyl)) +
  theme_spec("cell_press")

r <- fig_check(fixed, "cell_press", column = "single")
knitr::kable(
  r[r$status %in% c("pass", "fail"), c("check", "actual", "status")],
  row.names = FALSE
)
```

## Adjusting colour and shape

figspec separates publisher requirements from general guidance. If a plot
breaks a rule stated in the selected profile, the check fails. Other useful
findings are shown as guidance.

```{r}
knitr::kable(
  colour_safety_check(p, "cell_press")[, c("check", "actual", "status")],
  row.names = FALSE
)
```

Cell Press states that red and green should not be used together, so figspec
treats that combination as a failed requirement. The Royal Society publishes
in black and white by default, so figspec checks whether its colours remain
distinct in greyscale. When a publisher does not state either rule, figspec
still reports the relevant result but labels it `unspecified` instead of
presenting it as a requirement.

Colour does not have to carry all the information. Giving each series a
different point shape or line type provides a second way to identify it if two
colours look alike. figspec calls this *redundant coding* and reports whether
the plot provides that additional visual cue:

```{r}
knitr::kable(
  colour_safety_check(fixed, "royal_society")[, c("check", "actual", "status")],
  row.names = FALSE
)
```

`figspec_palettes()` lists the palettes that ship with the package, each with
its source. Cividis is safe both under colour vision deficiency and in
greyscale; Okabe-Ito is safe under colour vision deficiency but **not** in
greyscale, because two of its colours sit at almost the same lightness.

## Saving at the right size

`fig_save()` normally sizes the whole image, or canvas, to the width stated by
the journal. The plotting area inside that image is a second measurement, and
it can vary between figures even when their canvases are identical. The next
section introduces panel sizing, and the [panel-sizing guide](panels.html)
covers it in detail.

A common mistake is to save a figure much wider than the space available and
assume the publisher can reduce it safely. If a 180 mm figure is reduced to an
85 mm column, everything becomes 47% of its original size. An 8 pt label then
appears at only 3.8 pt.

```{r, eval = FALSE}
fig_save("figure_1.tiff", fixed, spec = "cell_press", column = "single")
```

`fig_save()` reads the required width and resolution from the registry, chooses
a graphics device for the requested file format, writes the file, and checks
the result—including whether that format is accepted.

You can see the available column widths before choosing one:

```{r}
fig_columns("science")
fig_columns("cell_press")
```

## Keep data panels aligned across a figure set

Saving every figure at the same width does not guarantee that the regions
containing the data will match. Axis labels, titles and legends use part of
the surrounding canvas. A figure with longer labels therefore has less room
left for its data, even when its image file is exactly the same width.

The two plots below use all 234 observations in `ggplot2::mpg`, with identical
variables, scales and a 150 mm canvas. Only the y-axis labels differ. Colour
and shape identify the drive type, so the groups remain distinguishable
without relying on colour alone.

### Before: equal canvases leave unequal room for data

```{r panel-before, fig.show = "hold", fig.width = 150/25.4, fig.height = 100/25.4, dpi = 200, out.width = "75%", fig.alt = c("A coloured fuel-economy plot with short y-axis labels and a relatively wide pale-blue data panel.", "The same coloured observations and scales with longer y-axis labels and a visibly narrower pale-blue data panel.")}
panel_colours <- setNames(
  figspec_palette("okabe_ito")[c(6, 7, 4)],
  c("4", "f", "r")
)
panel_shapes <- setNames(figspec_shapes(3), c("4", "f", "r"))

panel_base <- ggplot(
  ggplot2::mpg,
  aes(displ, hwy, colour = drv, shape = drv)
) +
  geom_point(alpha = 0.72, size = 1.8) +
  scale_colour_manual(values = panel_colours) +
  scale_shape_manual(values = panel_shapes) +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy",
       colour = "Drive type",
       shape = "Drive type") +
  theme_spec("cell_press", base_size = 8) +
  theme(
    legend.position = "bottom",
    panel.background = element_rect(
      fill = "#F3F8FC", colour = "#0072B2", linewidth = 0.7
    ),
    plot.background = element_rect(
      fill = "white", colour = "#9AA4AE", linewidth = 0.6
    )
  )

panel_figures <- list(
  short_labels = panel_base + labs(title = "Short tick labels"),
  long_labels = panel_base +
    scale_y_continuous(labels = function(x) paste(x, "miles per gallon")) +
    labs(title = "Long tick labels")
)

panel_figures$short_labels
panel_figures$long_labels
```

The longer labels leave less horizontal room for the data. `fig_panel_width()`
measures both layouts and finds the widest data panel that every figure can
share within the available 150 mm. `fig_panel_size()` then applies that exact
width and gives each figure the additional canvas its labels require.

### After: equal data panels with every label preserved

```{r panel-after, out.width = "75%", fig.alt = c("The coloured short-label plot after figspec applies the shared data-panel width.", "The coloured long-label plot with the same measured data-panel width and enough surrounding canvas to preserve its labels.")}
shared_width <- fig_panel_width(panel_figures, width = 150)
aligned_figures <- lapply(panel_figures, fig_panel_size, width = shared_width)
aligned_geometry <- lapply(aligned_figures, fig_geometry)

aligned_dir <- knitr::opts_current$get("fig.path")
dir.create(aligned_dir, recursive = TRUE, showWarnings = FALSE)
aligned_files <- file.path(
  aligned_dir, c("figspec-aligned-short.png", "figspec-aligned-long.png")
)
invisible(Map(function(figure, geometry, path) {
  ggplot2::ggsave(
    path, figure,
    width = geometry$canvas_width_mm,
    height = geometry$canvas_height_mm,
    units = "mm", dpi = 180, bg = "white"
  )
}, aligned_figures, aligned_geometry, aligned_files))
knitr::include_graphics(aligned_files)
```

Before adjustment, the two data panels could occupy
`r paste(round(attr(shared_width, "per_figure"), 1), collapse = " and ")` mm
respectively. After adjustment, both are exactly
`r round(shared_width, 1)` mm wide. The long-label figure receives a wider
canvas instead of a smaller plotting area, so its text remains readable and
the two data regions align.

This workflow is independent of journals. It is useful whenever figures must
align in a report, presentation, thesis, dashboard or composite layout. The
[panel-sizing guide](panels.html) explains the geometry and export options in
detail.

## Resolution requirements for different figure types

Publishers often require different resolutions for different kinds of
artwork. A colour or greyscale figure may need 300 dpi, while line art may need
600 to 1200 dpi. Twelve registry entries currently state a separate resolution
for line art.

In this context, line art means **monochrome artwork with no shading**.
Publishers use phrases such as "black and white graphic with no shading" and
"monochrome". Its sharp edges show stair-stepping easily, which is why it is
usually assigned the highest resolution.

A colour figure is therefore not line art and should not automatically be
checked against the line-art resolution. `fig_suggest_art_type()` examines what
the plot contains and recommends the appropriate category:

```{r}
line_art <- ggplot(ggplot2::mpg, aes(displ, hwy)) +
  geom_point(colour = "black")
fig_suggest_art_type(line_art, "bmj")
```

ggplot2's default bar fill is `#595959`, a mid grey, so a default bar chart is
grayscale art:

```{r}
fig_suggest_art_type(ggplot(ggplot2::mpg, aes(class)) + geom_bar())
```

`fig_check()` can classify a live plot from its layers and colours and then use
the matching resolution rule. A saved file may no longer reveal how the plot
was constructed, so `"auto"` uses the strictest recorded threshold rather than
risk approving a file with too little resolution.

## Checking a whole submission

One call can review a named collection of live plots. Here, both plots are
prepared for the quarterly report specification created earlier. Because the
plots are still editable R objects, figspec can inspect their text sizes,
colours and structure as well as their dimensions:

* The first plot shows how highway fuel economy changes with engine size.
  Colour and shape identify the number of cylinders.
* The second compares the distribution of highway fuel economy across vehicle
  classes. It is drawn horizontally so all seven class names remain readable.

They answer different questions and use different types of chart, just as the
figures in a real report often do. They are checked together because both must
follow the same report requirements—not because they are intended to be one
combined figure.

```{r submission-plots, fig.show = "hold", out.width = "100%", fig.alt = c("A fitted scatter plot of engine displacement and highway fuel economy. Colour and shape identify the number of cylinders.", "A horizontal box plot comparing highway fuel economy across vehicle classes, with each class shown in a different accessible colour.")}
submission_figures <- list(
  engine_size = ggplot(
    ggplot2::mpg,
    aes(displ, hwy, colour = factor(cyl), shape = factor(cyl))
  ) +
    geom_point(size = 2.5, alpha = 0.8, stroke = 0.7) +
    labs(
      title = "Engine size and highway fuel economy",
      x = "Engine displacement (litres)",
      y = "Highway miles per gallon",
      colour = "Cylinders",
      shape = "Cylinders"
    ) +
    fig_apply_spec(report_spec, base_size = 12),
  vehicle_class = ggplot(
    ggplot2::mpg,
    aes(hwy, reorder(class, hwy, median), fill = class)
  ) +
    geom_boxplot(
      width = 0.65,
      alpha = 0.9,
      outlier.shape = 21,
      outlier.size = 1.8,
      show.legend = FALSE
    ) +
    labs(
      title = "Highway fuel economy by vehicle class",
      x = "Highway miles per gallon",
      y = "Vehicle class"
    ) +
    fig_apply_spec(report_spec, base_size = 12)
)

submission_figures$engine_size
submission_figures$vehicle_class
```

```{r}
submission_check(submission_figures, report_spec, column = "full")
```

This first report checks everything that can be inspected while the plots are
still editable. Properties that belong to the finished files, such as the
saved format, resolution and file size, cannot be confirmed yet and remain
incomplete. After export, run `submission_check()` on the saved files to
verify those properties as well.

The same function accepts a directory or a vector of saved file paths. It can
check those files for dimensions, format, resolution and file size, but it can
no longer recover plot settings such as the original text sizes. For the most
complete assessment, check the editable plots first and the exported files
afterwards.

## In R Markdown or Quarto

Figures created in R Markdown or Quarto are often saved directly by knitr
rather than through `ggsave()`. Their dimensions and resolution therefore come
from chunk options such as `fig.width` and `dpi`.

```{r}
figspec_knitr_options("cell_press", "double")
```

Use `figspec_knitr_setup("cell_press", "double")` in the setup chunk to apply
the journal's size and resolution to every figure in the document.

The project specification created at the start works here too:

```{r}
figspec_knitr_options(report_spec, "full", height = 100, units = "mm")
```

## Reusing a figure set for another specification

A figure set may move to a new journal, report or production system. The new
requirements can be incompatible with the old ones: Cell Press wants type
between 6 and 8 pt, whereas PLOS ONE asks for 8 to 12 pt. Rather than redraw
the set by hand, re-export the plot objects against the new specification.

```{r, eval = FALSE}
fig_refit(my_plots, spec = "plos_one", output_dir = "figures_plos/")
```

`fig_refit()` applies the new theme and exports the figures again. It also
accepts a project specification instead of a registry name. The function needs
the original plot objects because a finished TIFF no longer contains reliable
information about settings such as the original text sizes.

## Preparing tables and other publication files

A publication may also include tables, graphical abstracts, video or audio,
each with requirements of its own. A publisher may specify how tables are
formatted, the dimensions of a graphical abstract, or the file types and
technical limits accepted for video and audio. figspec reports each kind of
requirement separately so it is clear which rules apply to which file.

```{r table-workflow}
table_report_spec <- list(
  name = "Quarterly outcomes report",
  tables = list(
    formats = c("html", "docx"),
    font_min_pt = 9,
    header_bold = TRUE,
    vertical_rules = FALSE
  )
)

summary_table <- table_apply_spec(
  mpg[1:10, c("manufacturer", "model", "displ", "hwy")],
  table_report_spec
)
table_check(summary_table, table_report_spec)
```

The same table can be exported and checked in one call:

```{r table-export}
table_path <- file.path(tempdir(), "vehicle-summary.html")
saved_table <- table_save(
  table_path,
  summary_table,
  table_report_spec,
  transform = FALSE
)
attr(saved_table, "figspec_table_report")
```

`table_apply_spec()` also accepts existing gt, flextable, kableExtra and grid
tables. Read [Build and check tables from R](tables.html) for worked examples
and the limits of each table system.

The lookup functions remain useful for publication assets with their own
requirements:

```{r other-assets}
graphical_abstract_spec("rsc")
table_spec("nature")
media_spec("science")
```

`media_check()` inspects an actual supplementary video or audio file. It
compares the file format, frame dimensions, file size, codec and audio bit rate
with the recorded requirements. Anything it cannot determine is identified
for manual review.

## When manual review is still needed

Some requirements cannot be checked automatically. If guidance gives
dimensions in pixels but no resolution for converting them, figspec keeps the
dimensions in pixels rather than guessing a physical size. If the guidance
does not state a rule, the report says that it is unspecified. To check text
size, provide the original plot because a saved raster image no longer records
the text's point size. The report lists each item that still needs manual
review. For tables, instructions about whether a title is concise, a footnote
is clear or every abbreviation is defined also remain for a person to assess.

figspec also measures differences between plotting-panel sizes without calling
them a failure. No publisher profile in the registry currently requires every
panel to have the same dimensions. You can use the measurements to align a set
when your report, presentation or figure layout needs it, without presenting a
design preference as a publisher requirement.

## Where to go next

* [Use figspec with different R plotting systems](figure-systems.html) for
  worked ggplot2, base R, lattice, grid and Plotly examples.
* [Learn about panel sizing](panels.html) to align the plotting areas across a
  set of figures.
* [Build and check tables from R](tables.html) to apply table requirements and
  verify exported files.
* [Browse the publication profiles](journals.html) and see how complete each
  registry entry is.
* [Review every function argument](options.html).
* Run `help(package = "figspec")` in an installed R session to browse the
  function index, organised by the job each function performs.
