---
title: "Meeting a journal's figure requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meeting a journal's figure requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(ggplot2)
library(figspec)
```

## The problem

Journals publish precise rules for the figures you submit, and the rules
differ. Cell Press asks for 85, 114 or 174 mm wide with type between 6 and 8 pt.
Science asks for 57, 121 or 184 mm with type no smaller than 5 pt. PNAS caps
type at 12 pt and asks that numerical axes reach zero. Nature wants panel
labels upright; the Royal Society wants them italicised.

Nothing in an ordinary plotting workflow makes any of that visible, and most
people find out at the production stage, after acceptance.

## The one moment that matters

**Type size only exists before you save.** Once a figure is a TIFF it is
pixels, and the point sizes are gone. That single fact shapes the whole
workflow: check the plot object, then save it.

```{r}
p <- ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_point() +
  labs(title = "Fuel economy", x = "Weight (1000 lbs)", y = "Miles per gallon",
       colour = "Cylinders")

check_journal(p, "cell_press", column = "single")
```

Two real failures in an ordinary ggplot, from defaults nobody thinks about.
ggplot2's base size is 11 pt, which puts text between 8.8 and 13.2 pt against
Cell Press's 6 to 8. And ggplot2's default palette contains a red and a green,
which Cell Press states should not be used together.

## Reading the report

Each requirement gets one of four outcomes, and only one of them is a problem
you must fix.

| Outcome | Meaning |
|---|---|
| `pass` | Meets the requirement |
| `fail` | Breaches it |
| `unspecified` | The publisher does not state this. Nothing can be concluded |
| `unknown` | The requirement exists, but this input cannot answer it |

`unspecified` splits further, and the distinction matters:

```{r}
r <- check_journal(p, "frontiers")
r[r$check %in% c("File size", "Height"), c("check", "requirement", "status")]
```

Frontiers' file-size rule was read and confirmed absent, so it is *not
specified by publisher*. Its height rule has not been resolved into a number by
anyone, so it is *not yet harvested for this journal*. The first is a fact
about Frontiers. The second is a fact about this registry, and saying so is the
difference between a registry you can trust and one that is confidently wrong.

## Fixing it

`theme_journal()` applies every requirement a theme can express: the font
family, the type floor and ceiling, stated line weights, and structural rules
such as Nature's requirement that axis lines and tick marks be drawn.

```{r}
fixed <- p +
  scale_colour_figspec("cividis") +
  scale_shape_manual(values = figspec_shapes(3)) +
  aes(shape = factor(cyl)) +
  theme_journal("cell_press")

r <- check_journal(fixed, "cell_press", column = "single")
r[r$status %in% c("pass", "fail"), c("check", "actual", "status")]
```

## Colour, and why shape matters

Three colour findings are reported, and which of them can *fail* depends on
what the publisher actually states.

```{r}
check_colour_safety(p, "cell_press")[, c("check", "actual", "status")]
```

Red-and-green is a **requirement** for Cell Press because Cell Press states it.
Greyscale separability is a requirement for the Royal Society, which prints in
black and white by default. Colour-vision safety is reported for every journal
but is always `unspecified`, because it is good practice rather than a
published rule.

Note the last row. If colours merge but shapes differ, the figure still reads:

```{r}
check_colour_safety(fixed, "royal_society")[
  , c("check", "actual", "status")]
```

`figspec_palettes()` lists the palettes that ship with the package, each with
its source. Cividis is safe both under colour vision deficiency and in
greyscale; Okabe-Ito is safe under colour vision deficiency but **not** in
greyscale, because two of its colours sit at almost the same lightness.

## Saving at the right size

This is the part that is easiest to get wrong. Type size is absolute. Save a
figure at 180 mm, have the journal place it in an 85 mm column, and everything
is scaled to 47%: a compliant 8 pt label arrives at 3.8 pt.

```{r, eval = FALSE}
ggsave_journal("figure_1.tiff", fixed, journal = "cell_press", column = "single")
```

`ggsave_journal()` takes the width from the registry, defaults the resolution
to the journal's stated minimum, picks a device that can render the required
font, and re-checks the file it wrote.

Column names come from the journal, not from a fixed vocabulary:

```{r}
fig_columns("science")
fig_columns("cell_press")
```

## Which resolution rule applies

Publishers set different resolutions for different kinds of artwork, and the
gap is large: 300 dpi generally, against 600 to 1200 dpi for line art. Twelve
entries in the registry state a separate line-art figure.

Most R plots are line art. PNAS says so explicitly, giving "line art, e.g., bar
graphs". So a bar chart saved at 300 dpi can be a quarter of what the journal
asked for, and nothing in an ordinary workflow says so.

```{r}
suggest_art_type(ggplot(mtcars, aes(factor(cyl))) + geom_bar(), "bmj")
```

`check_journal()` checks against the general minimum unless you say otherwise,
and says so when a plot looks like line art and the journal holds line art to
more. It does not change the verdict for you: publishers genuinely disagree
about whether a coloured plot counts, so the choice stays yours.

## Checking a whole submission

```{r, eval = FALSE}
check_submission("figures/", "cell_press")
#> ✔ Figure_1.tiff  single  all requirements met
#> ✖ Figure_2.tiff  double  failed: Resolution, File format
```

Files are checked as files, so geometry, format, resolution and file size can
be judged but type size cannot. That is why step one happens before saving.

## In R Markdown or Quarto

A large share of academic figures never pass through `ggsave()` at all: they
come out of a knitr chunk at whatever `fig.width` happens to be.

```{r}
figspec_chunk_opts("cell_press", "double")
```

Put that in a setup chunk with `figspec_knitr_setup("cell_press", "double")`
and every figure in the document is emitted at the journal's size.

## When the paper is rejected

The next journal has different rules, and they can be incompatible: Cell Press
wants type between 6 and 8 pt, PLOS ONE between 8 and 12.

```{r, eval = FALSE}
refit_journal(my_plots, journal = "plos_one", outdir = "figures_plos/")
```

`refit_journal()` re-themes and re-exports the set. It works from plot objects
and refuses saved files, because type size cannot be recovered from a finished
TIFF and a re-fitted file could not be trusted.

## Other display items

Figures are not the only thing with rules.

```{r}
graphical_abstract_spec("rsc")
```

`table_spec()` and `media_spec()` do the same for tables and for supplementary
video and audio.

## What figspec will not tell you

It will not tell you a figure is fine when the publisher never said what fine
means. It will not convert a stated pixel size into millimetres without a
stated resolution. It will not grade a recommendation as a failure. And it will
not recover type size from a saved raster.

Those are all cases where the honest answer is "check this yourself", and
figspec says so rather than guessing.
