---
title: "Use figspec with different R plotting systems"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use figspec with different R plotting systems}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7.2,
  fig.height = 4.8,
  dpi = 160
)
library(figspec)

embed_saved_figure <- function(path, alt, width = "78%") {
  uri <- knitr::image_uri(path)
  knitr::asis_output(sprintf(
    paste0(
      '<p class="figspec-embedded-plot">',
      '<img src="%s" style="width:%s;max-width:100%%;height:auto" ',
      'alt="%s"></p>'
    ),
    uri, width, alt
  ))
}

summarise_export <- function(path) {
  geometry <- fig_geometry(path)
  data.frame(
    plotting_system = attr(path, "figspec_system"),
    requirements_applied = attr(path, "figspec_transformation"),
    canvas_mm = sprintf(
      "%.0f × %.0f",
      geometry$canvas_width_mm,
      geometry$canvas_height_mm
    ),
    stringsAsFactors = FALSE
  )
}

```

figspec does not require every figure to begin in ggplot2. You can give
`fig_save()` a ggplot2 plot, a lattice plot, a grid graphic, a Plotly widget or
base R plotting code. The same specification then controls the output size,
resolution and format. figspec also applies the requirements and visual
defaults that the plotting system makes available before writing the file.

The important difference is *when* figspec receives the figure. An editable
plot object or a base R drawing function still contains instructions that can
be changed. A completed PNG or a recorded base plot has already lost much of
that information. figspec can verify the finished file, but it cannot safely
reconstruct every original colour, type or line setting.

## Use one specification with several plotting systems

The examples below use real observations from `ggplot2::mpg` and the same
project specification. The report requires a 110 × 75 mm PNG at 160 dpi, text
between 9 and 11 points, and lines at least 0.6 points wide:

```{r shared-specification}
figure_spec <- list(
  name = "Shared research report",
  source_url = "internal:research-report-guide",
  verified_on = as.character(Sys.Date()),
  columns = list(full = 110),
  formats = "png",
  dpi_min = 160,
  font_min_pt = 9,
  font_max_pt = 11,
  min_line_pt = 0.6,
  axis_lines_and_ticks = TRUE
)

vehicle_data <- ggplot2::mpg
```

The specification is an ordinary named R list. A registry profile such as
`"cell_press"` can be supplied in exactly the same position.

## ggplot2 and patchwork

For a ggplot2 figure, figspec can work with its layers, scales and theme. It
can apply the recorded typography and axis requirements, add accessible
colour and shape scales, control the data-panel dimensions and inspect the
editable plot as well as the completed file.

```{r ggplot-example}
ggplot_figure <- ggplot2::ggplot(
  vehicle_data,
  ggplot2::aes(displ, hwy, colour = drv, shape = drv)
) +
  ggplot2::geom_point(size = 1.8, alpha = 0.75) +
  ggplot2::labs(
    title = "Engine size and highway fuel economy",
    x = "Engine displacement (litres)",
    y = "Highway fuel economy",
    colour = "Drive type",
    shape = "Drive type"
  )

ggplot_path <- file.path(tempdir(), "figspec-system-ggplot.png")
ggplot_saved <- fig_save(
  ggplot_path,
  ggplot_figure,
  spec = figure_spec,
  column = "full",
  height = 75,
  check = FALSE
)
summarise_export(ggplot_saved)
```

```{r show-ggplot-export, echo=FALSE, results='asis'}
embed_saved_figure(
  ggplot_saved,
  paste(
    "A scatter plot made with ggplot2 showing engine displacement and",
    "highway fuel economy. Colour and shape distinguish drive types."
  )
)
```

Patchwork compositions use the same `fig_save()` call. figspec applies the
typography requirements across the component plots and can set the size of
each data panel. Patchwork continues to control the arrangement.

## Base R graphics

Base R normally draws directly onto the active graphics device, so there may
be no plot object to hand to another function. Give figspec the drawing code
inside `function() { ... }` or a one-sided formula. figspec can then open the
correct device first and set its type size, font, line-width and colour-palette
defaults before the drawing runs.

```{r base-example}
base_figure <- function() {
  drive <- factor(vehicle_data$drv)
  graphics::plot(
    vehicle_data$displ,
    vehicle_data$hwy,
    col = as.integer(drive),
    pch = c(16, 17, 15)[drive],
    xlab = "Engine displacement (litres)",
    ylab = "Highway fuel economy",
    main = "Engine size and highway fuel economy"
  )
  graphics::legend(
    "topright",
    legend = levels(drive),
    col = seq_along(levels(drive)),
    pch = c(16, 17, 15),
    title = "Drive type",
    bty = "n"
  )
}

base_path <- file.path(tempdir(), "figspec-system-base.png")
base_saved <- fig_save(
  base_path,
  base_figure,
  spec = figure_spec,
  column = "full",
  height = 75,
  check = FALSE
)
summarise_export(base_saved)
```

```{r show-base-export, echo=FALSE, results='asis'}
embed_saved_figure(
  base_saved,
  paste(
    "A scatter plot made with base R showing engine displacement and",
    "highway fuel economy, exported through figspec."
  )
)
```

The distinction between code and a recorded plot matters. Explicit settings
inside the drawing function still take precedence: if the code asks for
2-point text or a particular colour, figspec does not rewrite the R commands.
Pass defaults through the function when you want figspec to choose them. A
`recordedplot` can be exported and checked, but it cannot be reliably restyled;
use the drawing function whenever the original code is available.

## Lattice

A lattice plot is an editable `trellis` object. figspec supplies specification-
aware typography, colours, symbols, line types and line widths through lattice
settings, then prints the transformed object on the requested device.

```{r lattice-example}
if (requireNamespace("lattice", quietly = TRUE)) {
  lattice_figure <- lattice::xyplot(
    hwy ~ displ,
    data = vehicle_data,
    groups = drv,
    auto.key = list(space = "right", title = "Drive type"),
    xlab = "Engine displacement (litres)",
    ylab = "Highway fuel economy",
    main = "Engine size and highway fuel economy"
  )

  lattice_path <- file.path(tempdir(), "figspec-system-lattice.png")
  lattice_saved <- fig_save(
    lattice_path,
    lattice_figure,
    spec = figure_spec,
    column = "full",
    height = 75,
    check = FALSE
  )
  summarise_export(lattice_saved)
}
```

```{r show-lattice-export, echo=FALSE, results='asis'}
if (exists("lattice_saved")) {
  embed_saved_figure(
    lattice_saved,
    paste(
      "A lattice scatter plot of engine displacement and highway fuel",
      "economy, with drive types distinguished by colour and symbol."
    )
  )
}
```

## Grid graphics

Grid figures are trees of graphical objects. figspec can give the tree
specification-aware font and line defaults. Settings written directly on a
child grob remain unchanged because they are explicit design choices, not
defaults that figspec can safely replace.

```{r grid-example}
normalise <- function(x) {
  (x - min(x)) / diff(range(x))
}

grid_figure <- grid::grobTree(
  grid::rectGrob(
    x = 0.55, y = 0.5, width = 0.78, height = 0.72,
    gp = grid::gpar(fill = "#F3F8FC", col = "#1A7391")
  ),
  grid::pointsGrob(
    x = grid::unit(0.18 + 0.68 * normalise(vehicle_data$displ), "npc"),
    y = grid::unit(0.14 + 0.66 * normalise(vehicle_data$hwy), "npc"),
    pch = 21,
    size = grid::unit(2.1, "mm"),
    gp = grid::gpar(fill = "#56B4E9", col = "white")
  ),
  grid::textGrob(
    "Engine size and highway fuel economy",
    x = 0.55, y = 0.91
  ),
  grid::textGrob("Engine displacement", x = 0.55, y = 0.06),
  grid::textGrob(
    "Highway fuel economy",
    x = 0.08, y = 0.5, rot = 90
  )
)

grid_path <- file.path(tempdir(), "figspec-system-grid.png")
grid_saved <- fig_save(
  grid_path,
  grid_figure,
  spec = figure_spec,
  column = "full",
  height = 75,
  check = FALSE
)
summarise_export(grid_saved)
```

```{r show-grid-export, echo=FALSE, results='asis'}
embed_saved_figure(
  grid_saved,
  paste(
    "A grid graphic using real vehicle data, with a pale blue data area",
    "and specification-aware text and line defaults."
  )
)
```

Grid does not expose a universal concept of a ggplot2 data panel, so exact
panel sizing is not offered for a general grid tree. Set the complete canvas
with `width` and `height`.

## Plotly and other HTML widgets

Plotly stores editable layout and trace definitions. figspec can apply fonts,
axis lines, colours, marker symbols, line types and line weights before export.
For a PNG, JPEG or TIFF, it renders the widget in a local Chrome or Chromium
window at the requested pixel dimensions. Nothing is uploaded to an external
service.

```{r plotly-code, eval=FALSE}
plotly_figure <- plotly::plot_ly(
  vehicle_data,
  x = ~displ,
  y = ~hwy,
  color = ~drv,
  symbol = ~drv,
  type = "scatter",
  mode = "markers"
)

plotly_saved <- fig_save(
  "plotly-figure.png",
  plotly_figure,
  spec = figure_spec,
  column = "full",
  height = 75
)
```

The website build runs that workflow against a real local browser:

```{r plotly-browser-export, echo=FALSE, results='asis', eval=identical(Sys.getenv("FIGSPEC_BUILD_BROWSER_EXAMPLES"), "true")}
if (
  requireNamespace("plotly", quietly = TRUE) &&
  requireNamespace("htmlwidgets", quietly = TRUE) &&
  requireNamespace("webshot2", quietly = TRUE)
) {
  plotly_figure <- plotly::plot_ly(
    vehicle_data,
    x = ~displ,
    y = ~hwy,
    color = ~drv,
    symbol = ~drv,
    type = "scatter",
    mode = "markers"
  )
  plotly_path <- file.path(tempdir(), "figspec-system-plotly.png")
  plotly_saved <- suppressMessages(fig_save(
    plotly_path,
    plotly_figure,
    spec = figure_spec,
    column = "full",
    height = 75,
    check = FALSE
  ))
  embed_saved_figure(
    plotly_saved,
    paste(
      "A browser-rendered Plotly scatter plot using the same vehicle data",
      "and project specification as the other examples."
    )
  )
}
```

Plotly's native SVG and PDF export uses its optional Kaleido environment.
Other HTML widgets can be exported to PNG, JPEG or TIFF at exact dimensions,
but figspec does not assume that an unfamiliar widget exposes editable fonts,
lines or data marks.

## Verify the completed files

The plotting systems differ, but the files they produce can be checked in one
way. This example verifies the dimensions, resolution and format of every file
created above:

```{r verify-exports}
exports <- list(
  ggplot2 = ggplot_saved,
  base_R = base_saved,
  grid = grid_saved
)
if (exists("lattice_saved")) exports$lattice <- lattice_saved
if (exists("plotly_saved")) exports$plotly <- plotly_saved

verification <- lapply(exports, function(path) {
  report <- fig_check(path, figure_spec, column = "full")
  report[
    report$check %in% c("Width", "Resolution", "File format"),
    c("check", "actual", "status")
  ]
})
verification
```

A raster image does not retain the original point size or the identity of its
plotting system. Check the editable object when you need evidence about its
type, layers or mappings, and check the exported file for the properties that
only exist after export.

## Choose the right input

| Figure you have | What to give figspec | What figspec can apply before export | Important limit |
|---|---|---|---|
| ggplot2 | The plot object | Theme, type, axes, colours, shapes and line rules | Most complete inspection and exact panel sizing |
| patchwork | The composition | Typography across components | Patchwork controls arrangement; exact panel sizing is available |
| base R | A function or one-sided formula containing the drawing code | Device font, type, line and palette defaults | Explicit drawing arguments take precedence |
| lattice | The `trellis` object | Type, colours, symbols, line types and line widths | Exact panel sizing is not available |
| grid | A grob or grob tree | Inherited font and line defaults | Explicit child settings remain unchanged; no general panel model |
| Plotly | The Plotly object | Layout fonts and axes; trace colours, markers and lines | Raster export needs a local browser; vector export needs Kaleido |
| Another HTML widget | The widget object | Exact raster dimensions | No general restyling contract |
| Recorded base plot | The `recordedplot` | Nothing safely | It can still be exported and checked |
| Finished file | Its file path | Nothing—the drawing is complete | figspec verifies what the file still records |

If exact data-panel dimensions or detailed layer checks are central to the
work, use a ggplot2-compatible figure. If the figure already belongs in
another plotting system, figspec still provides one repeatable route from the
editable input to a correctly sized and verified file without requiring you
to rebuild it in ggplot2.
