---
title: "Which journal am I submitting to?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Which journal am I submitting to?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
```

```{r setup, include = FALSE}
library(figspec)
fmt <- function(x, unit = "") {
  if (is.null(x) || !length(x) || all(is.na(x))) return("–")
  paste0(paste(unlist(x), collapse = ", "), unit)
}
```

Everything below is generated from the registry itself, so it cannot drift from
what the package actually holds. A dash means the publisher states nothing, or
nobody has harvested it yet — `journal_spec()` tells you which.

```{r}
reg <- figspec:::load_registry()
rows <- do.call(rbind, lapply(reg, function(j) {
  cols <- if (!is.null(j$columns)) {
    paste(paste0(names(j$columns), " ", unlist(j$columns)), collapse = ", ")
  } else if (!is.null(j$width_min_mm) || !is.null(j$width_max_mm)) {
    paste0(j$width_min_mm %||% "?", "–", j$width_max_mm %||% "?")
  } else "–"
  data.frame(
    Journal = j$name,
    id = j$id,
    `Width (mm)` = cols,
    `dpi` = fmt(j$dpi_min),
    `Line art` = fmt(j$dpi_line_art),
    `Type (pt)` = if (is.null(j$font_min_pt)) "–" else
      paste0(j$font_min_pt, if (!is.null(j$font_max_pt)) paste0("–", j$font_max_pt) else "+"),
    Formats = fmt(toupper(unlist(j$formats %||% NULL))),
    check.names = FALSE, stringsAsFactors = FALSE)
}))
rows <- rows[order(rows$Journal), ]
knitr::kable(rows, row.names = FALSE)
```

## Where publishers disagree

The registry makes some conflicts plain, and they are the reason a single
"industry standard" does not exist.

**Panel labels.** Cell Press wants capital letters. AGU, Nature, Springer and
the Royal Society all want lower case. Nature wants them *upright, not italic*;
the Royal Society wants them *italicised*. Same element, opposite rules.

**Outlining text.** BMJ requires text in EPS files to be outlined. Nature
instructs authors not to outline text.

**Line art resolution.** Twelve entries state a separate figure for line art,
from 600 dpi (IEEE, Wiley) to 1200 dpi (BMJ, ACS, OUP, Springer, Taylor &
Francis) — against a general minimum of 300 almost everywhere. See
`suggest_art_type()`, and note that line art means *monochrome*: five
publishers define it as black and white.

**Colour in print.** Frontiers and PLOS say nothing about greyscale. The Royal
Society, Sage and Taylor & Francis print in black and white by default, so
figures must survive it. RSC prints colour free. RSC's *books* make it depend
on the author's contract.

## How complete is each entry

```{r}
st <- registry_status()
st <- st[order(-st$stated), c("id", "stated", "confirmed_absent", "unharvested", "verified_on")]
names(st) <- c("id", "stated", "confirmed absent", "not yet harvested", "read on")
knitr::kable(st, row.names = FALSE)
```

`stated` is what the publisher says. `confirmed absent` is what somebody read
the page for and found missing. `not yet harvested` is what nobody has checked.
The three always sum to the full field list, so a half-finished entry cannot
pass as a complete one.
