---
title: "Find and use publication specifications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Find and use publication specifications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, include = FALSE}
library(figspec)
library(ggplot2)

fmt <- function(x, unit = "") {
  if (is.null(x) || !length(x) || all(is.na(x))) return("–")
  paste0(paste(unlist(x), collapse = ", "), unit)
}
```

figspec includes a registry of carefully sourced figure requirements from
major publishers and selected journals. It currently contains 29 profiles:
22 record publisher-wide guidance that applies across large journal
portfolios, while seven cover individual journals with their own requirements.

Each recorded requirement includes the page from which it was taken and the
date on which that page was reviewed. A missing value is never treated as
permission or as a passing result: figspec distinguishes a requirement that
the source does not state from one that has not yet been reviewed for that
profile.

The registry is optional. You can also supply requirements maintained by your
project or organisation, or load a registry file maintained by your team. A
house style is different: it controls visual choices, while a specification
defines requirements that can be checked.

Figure and table guidance is reviewed separately. Some bundled profiles
already contain table requirements, while others have not yet had their table
guidance reviewed. `registry_status()` shows both kinds of coverage. The full
table workflow can also use requirements supplied by a project or organisation;
see [Build and check tables from R](tables.html).

## Find a relevant profile

`spec_list()` lists the available publisher and journal profiles. Use the
registry ID in functions such as `spec_get()`, `fig_apply_spec()`,
`fig_save()` and `fig_check()`. You can list everything or narrow the results
by discipline:

```{r find-profiles}
physics_profiles <- spec_list(discipline = "physics")
knitr::kable(
  physics_profiles[, c(
    "id", "name", "publisher", "single_mm", "double_mm", "verified_on"
  )],
  col.names = c(
    "Registry ID", "Profile", "Publisher", "Single (mm)", "Double (mm)",
    "Verified"
  ),
  row.names = FALSE
)
```

A publisher-wide profile is appropriate when a journal follows that
publisher's central figure guidance. Use a journal-specific profile when the
journal publishes additional or different requirements. Always follow any
instructions supplied directly by the journal for the article you are
preparing.

## Inspect a profile before using it

`spec_get()` presents the recorded requirements, source and verification
date together. `fig_columns()` extracts the available figure widths when that
is all you need:

```{r inspect-profile}
cell_press <- spec_get("cell_press")
cell_press
fig_columns(cell_press)
```

The returned specification is an ordinary R object. Pass it directly to the
rest of the figspec workflow, or use the shorter registry ID. Here the selected
profile is applied to all 234 observations in `ggplot2::mpg`:

```{r apply-profile, fig.width = 85/25.4, fig.height = 70/25.4, dpi = 200, out.width = "70%", fig.alt = "A coloured scatter plot of engine displacement and highway fuel economy after applying the Cell Press profile."}
profile_plot <- ggplot(
  ggplot2::mpg,
  aes(displ, hwy, colour = factor(cyl), shape = factor(cyl))
) +
  geom_point(size = 1.8, alpha = 0.72) +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy",
       colour = "Cylinders",
       shape = "Cylinders") +
  fig_apply_spec(cell_press)

profile_plot
```

Because the plot is still editable, `fig_check()` can assess its text, colour,
shapes and layout as well as its intended width:

```{r check-profile-plot}
profile_result <- fig_check(profile_plot, cell_press, column = "single")
knitr::kable(
  profile_result[
    profile_result$check %in% c(
      "Width", "Type size", "Colour pairs", "Colour vision",
      "Redundant coding"
    ),
    c("check", "actual", "status")
  ],
  row.names = FALSE
)
```

## Browse all included profiles

The compact table below is rebuilt from `spec_list()` and `spec_get()`
whenever the website is generated. A dash keeps the overview readable; inspect
the profile to learn whether its source is silent on that requirement or the
field has not yet been reviewed.

```{r all-profiles, echo = FALSE, results = "asis"}
ids <- spec_list()$id
reg <- stats::setNames(lapply(ids, spec_get), ids)
rows <- do.call(rbind, lapply(reg, function(j) {
  cols <- if (!is.null(j$columns) && length(j$columns)) {
    paste(paste0(names(j$columns), " ", unlist(j$columns)), collapse = ", ")
  } else if (!is.null(j$width_min_mm) || !is.null(j$width_max_mm)) {
    lower <- if (is.null(j$width_min_mm)) "?" else j$width_min_mm
    upper <- if (is.null(j$width_max_mm)) "?" else j$width_max_mm
    paste0(lower, "–", upper)
  } else {
    "–"
  }
  data.frame(
    Profile = j$name,
    `Registry ID` = j$id,
    `Figure width (mm)` = cols,
    `General dpi` = fmt(j$dpi_min),
    `Line-art dpi` = fmt(j$dpi_line_art),
    `Text size (pt)` = if (is.null(j$font_min_pt)) {
      "–"
    } else {
      paste0(
        j$font_min_pt,
        if (!is.null(j$font_max_pt)) paste0("–", j$font_max_pt) else "+"
      )
    },
    `Accepted formats` = fmt(toupper(unlist(j$formats))),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}))
rows <- rows[order(rows$Profile), ]

cat(sprintf(
  '<details class="figspec-details"><summary>View all %d publication profiles</summary>\n',
  nrow(rows)
))
cat(knitr::kable(rows, format = "html", row.names = FALSE), sep = "\n")
cat("\n</details>")
```

## How publication requirements differ

Publishers prepare figures for different page layouts, printing processes and
digital formats, so their requirements are not identical. A figure that meets
one publication's guidance may therefore need changes before it is submitted
elsewhere. The registry records differences such as these:

**Panel labels.** Cell Press asks for capital letters. AGU, Nature, Springer
and the Royal Society use lower-case letters. Nature specifies upright labels,
whereas the Royal Society specifies italics.

**Text in EPS files.** BMJ asks authors to outline text in EPS artwork. Nature
asks authors not to outline it.

**Line-art resolution.** Twelve profiles record a separate resolution for line
art. The recorded thresholds range from 600 dpi for IEEE and Wiley to 1200 dpi
for BMJ, ACS, OUP, Springer and Taylor & Francis, compared with a common
general minimum of 300 dpi. Where sources define line art, they commonly
describe monochrome artwork without shading. `fig_suggest_art_type()` helps select
the applicable resolution rule.

**Colour in print.** Some publications accept or provide colour, while others
expect figures to remain readable when printed in greyscale. The Royal Society,
Sage and Taylor & Francis record black-and-white printing as the default; RSC
records free colour for journals, while its book policy depends on the
author's contract.

These are production choices, not disagreements. figspec keeps them separate
so that the requirements applied to a figure match its intended destination.

## Check what each profile covers

`registry_status()` distinguishes three kinds of registry coverage:

* **Stated** means the source gives a requirement for that field.
* **Confirmed absent** means the source was reviewed but gives no requirement
  for that field.
* **Not yet harvested** means the field has not yet been reviewed or entered
  for that profile.

The categories account for every field tracked by the registry. Keeping them
separate prevents an unreviewed field from being mistaken for a requirement
that the publisher does not state.

```{r coverage, echo = FALSE, results = "asis"}
st <- registry_status()
st <- st[
  order(-st$stated),
  c("id", "stated", "confirmed_absent", "unharvested", "verified_on")
]
names(st) <- c(
  "Registry ID", "Stated", "Confirmed absent", "Not yet harvested", "Verified"
)

cat(sprintf(
  '<details class="figspec-details"><summary>View coverage for all %d profiles</summary>\n',
  nrow(st)
))
cat(knitr::kable(st, format = "html", row.names = FALSE), sep = "\n")
cat("\n</details>")
```

## Use your own specification

Publication profiles are only one source of requirements. The same functions
accept a named list describing a report, thesis, presentation, organisation or
production workflow. This example defines a 160 mm quarterly-report figure,
builds a real plot from `ggplot2::mpg`, and applies the required text size:

```{r custom-specification, fig.width = 160/25.4, fig.height = 105/25.4, dpi = 150, out.width = "75%", fig.alt = "A coloured scatter plot of engine displacement and highway fuel economy prepared for a quarterly report specification."}
project_spec <- spec_get(list(
  name = "Quarterly outcomes report",
  columns = list(full = 160),
  formats = c("png", "pdf"),
  dpi_min = 300,
  font_min_pt = 9,
  min_line_pt = 0.5
))

project_colours <- setNames(
  figspec_palette("okabe_ito")[c(6, 7, 4)],
  c("4", "f", "r")
)

project_plot <- ggplot(
  ggplot2::mpg,
  aes(displ, hwy, colour = drv, shape = drv)
) +
  geom_point(size = 2, alpha = 0.72) +
  scale_colour_manual(values = project_colours) +
  scale_shape_figspec() +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy",
       colour = "Drive type",
       shape = "Drive type") +
  theme_spec(project_spec, base_size = 9)

project_plot
```

The example below writes an actual 300 dpi PNG, reopens it and checks its saved
properties against the project specification:

```{r check-custom-export}
project_file <- tempfile(fileext = ".png")
fig_save(
  project_file,
  project_plot,
  spec = project_spec,
  column = "full",
  dpi = 300,
  check = FALSE
)

project_result <- fig_check(project_file, project_spec)
knitr::kable(
  project_result[
    project_result$check %in% c(
      "File validity", "Width", "Resolution", "File format"
    ),
    c("check", "requirement", "actual", "status")
  ],
  row.names = FALSE
)
```

```{r cleanup-custom-export, include = FALSE}
unlink(project_file)
```

A field omitted from your own specification is reported as **unspecified**. It
is not called *not yet harvested*, because you supplied the requirements and
the omission is not a gap in figspec's review of a publication source.

Pass no specification at all and `fig_check()` becomes an inspection: it
reports the figure's size, resolution, format and file size without assigning
a pass or fail where no requirement was provided.
