---
title: "Sizing figures by panel, not by canvas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sizing figures by panel, not by canvas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Every figure has two widths, and the one you can normally set is not the one a
reader notices.

* The **canvas** is the image file. `ggsave(width = 85, units = "mm")` sets this.
* The **panel** is the plot area left over once the axis title, the tick
  labels, the tick marks, the legend and the margins have taken their share.

You set the canvas. The panel gets whatever is left. So a figure with long tick
labels has a smaller plot area than one with short labels, even when both were
saved at exactly the same width — and side by side on a page, they look
uneven. That mismatch is the most common reason a set of figures looks
unfinished, and it is invisible in the code.

figspec lets you set the panel instead, and works the canvas out for you.

```{r setup}
library(figspec)
library(ggplot2)
```

## The problem, in three figures

Three plots of the same data. They differ only in the magnitude of the y
values, which changes how wide the tick labels are.

```{r}
figs <- list(
  fig1 = ggplot(mtcars, aes(wt, mpg)) + geom_point(),
  fig2 = ggplot(mtcars, aes(wt, mpg * 100000)) + geom_point(),
  fig3 = ggplot(mtcars, aes(wt, mpg / 100)) + geom_point()
)
```

Saved the usual way, all three are 85 mm wide — the single-column width Cell
Press asks for. `check_submission()` measures what that leaves for the plot
area:

```{r}
check_submission(figs, "cell_press")
```

Same canvas, three different plot areas. Nothing here is wrong: every figure
meets the journal's stated width. But the middle one has millimetres less room
to draw in, and on the printed page you can see it.

Note how that is reported. **No publisher states a rule about panel
consistency, so figspec never calls it a failure.** It is an observation about
your figures, printed apart from the pass and fail lines and never counted
among them. A green tick against an invented rule would be the same error as
inventing the rule.

## Setting a panel width

`fig_save()` takes `panel_width`. Give it a number and the plot area comes out
that size exactly, whatever the labels do:

```{r}
f <- tempfile(fileext = ".png")

fig_geometry(fig_save(f, figs$fig1, panel_width = 62))
fig_geometry(fig_save(f, figs$fig2, panel_width = 62))
```

Both panels are 62 mm. The canvases differ, because the second figure needs
more room for its labels — which is the right way round. The panel is what you
care about; the canvas is the consequence.

## Making a set match

For a set, the panel width you can share is set by the figure with the *least*
room — the one with the longest tick labels. `fig_panel_width()` measures every
figure and returns the widest panel all of them can use:

```{r}
pw <- fig_panel_width(figs, journal = "cell_press", format = "tiff")
pw
attr(pw, "per_figure")
```

`fig2` is the binding constraint. Pass that number back to `fig_save()` and
every figure comes out with the same plot area, still at the journal's 85 mm
canvas — the slack goes to the margin rather than stretching the panel:

```{r}
for (nm in names(figs)) {
  fig_save(file.path(tempdir(), paste0(nm, ".tiff")), figs[[nm]],
           journal = "cell_press", panel_width = pw)
}

check_submission(lapply(figs, fig_panel_size, width = pw), "cell_press")
```

## Letting the journal decide

If you only have one figure and simply want it to use its column as fully as
possible, `panel_width = "max"` takes the widest panel that still fits:

```{r}
fig_geometry(fig_save(f, figs$fig1, journal = "cell_press", panel_width = "max"))
```

Note that `"max"` is per figure. It gives *this* figure the largest panel it
can have, which is not the same as giving a *set* matching panels — for that
you want `fig_panel_width()` above.

## Canvas and panel together

`width` and `panel_width` are not alternatives. They are two constraints on one
equation:

```
canvas = panel + decoration
```

The third term is measured, not guessed, so figspec can solve for whichever you
leave out. Give both and both are honoured — the leftover becomes margin:

```{r}
fig_geometry(fig_save(f, figs$fig1, width = 120, panel_width = 62))
```

Ask for a pair that cannot exist and it refuses, naming the two values that
would work. You cannot arrive at those without doing the measurement, so the
error carries them:

```{r, error = TRUE}
fig_save(f, figs$fig1, width = 70, panel_width = 62)
```

Under a journal, the column width pins the canvas — a figure that is not the
column width gets rescaled in production, and rescaling is what drives type
below the stated minimum. So an over-wide panel is refused for the same reason:

```{r, error = TRUE}
fig_save(f, figs$fig1, journal = "cell_press", panel_width = 200)
```

## Facets and compositions

A panel width applies to *each* panel, which is what keeps panels comparable
between figures rather than between layouts. A three-facet figure at 30 mm
panels is three 30 mm panels wide, not 30 mm in total:

```{r}
faceted <- ggplot(mtcars, aes(wt, mpg)) + geom_point() + facet_wrap(~cyl)
fig_geometry(fig_save(f, faceted, panel_width = 30))
```

The same holds for a patchwork composition, measured through the nesting.

## Setting a panel without saving

`fig_panel_size()` does the geometry and hands back a figure you can print,
inspect or pass on. It keeps the plot it was built from, so the result can
still be checked in full — type size, font and colour live on the plot, not on
the layout, and would otherwise be lost:

```{r}
sized <- fig_panel_size(figs$fig1, width = 62, height = 45)
fig_geometry(sized)
```

## What a figure is right now

`fig_geometry()` reports where the space in a figure is going, which is usually
the fastest way to see why one is not the size you expected:

```{r}
fig_geometry(figs$fig2)
```

An unsized plot reports `NA` for its panel and canvas, and a real number for
its decoration. That is not a failure to measure — a panel with no size of its
own genuinely has no width until a canvas is chosen. The decoration, which
depends only on the labels and the font, is known either way.

The per-side figures are usually what you want. A total tells you the canvas
has to grow; the sides tell you *why*, and they have different fixes — a wide
right margin is a legend to move, a wide left margin is axis labels to shorten.

```{r fig.width = 5, fig.height = 3.6}
plot(fig_geometry(fig_panel_size(figs$fig1, width = 60, height = 40)))
```

```{r, include = FALSE}
unlink(f)
unlink(file.path(tempdir(), paste0(names(figs), ".tiff")))
```
