---
title: "Control figure and panel dimensions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Control figure and panel dimensions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

embed_plot <- function(draw, width, height, dpi, alt, out_width = "100%") {
  path <- tempfile(fileext = ".png")
  if (requireNamespace("ragg", quietly = TRUE)) {
    ragg::agg_png(
      path, width = width, height = height, units = "in", res = dpi,
      background = "white"
    )
  } else {
    grDevices::png(
      path, width = width, height = height, units = "in", res = dpi,
      bg = "white"
    )
  }
  tryCatch(draw(), finally = grDevices::dev.off())
  uri <- knitr::image_uri(path)
  unlink(path)
  knitr::asis_output(sprintf(
    paste0(
      '<p class="figspec-embedded-plot">',
      '<img src="%s" style="width:%s;max-width:100%%;height:auto" ',
      'alt="%s"></p>'
    ),
    uri, out_width, alt
  ))
}
```

```{r setup}
library(figspec)
library(ggplot2)
```

The **plot panel** is the rectangle in which the data are drawn. Panel sizing
lets you set the physical width or height of that rectangle directly. figspec
then calculates the complete image needed around it.

This matters when several figures will appear together in a paper, report or
presentation. Readers compare the axes and data regions, so those regions
should line up even when one figure has longer labels or a larger legend.

Two measurements determine the result:

* The **canvas** is the complete image, including labels, legends and margins.
  A call such as `ggsave(width = 85, units = "mm")` sets the canvas width.
* The **panel** is the area in which the data are drawn. It receives whatever
  space remains after the axes, labels, legend and margins have been placed.

Labels, legends and margins have fixed physical sizes. On canvases of equal
width, a longer label leaves less room for the data. figspec holds the panel
width constant and calculates how much canvas each figure needs around it.

## See where the space around the panel goes

`fig_geometry()` shows how a figure's canvas is divided between its data panel
and the material around it. This example uses all 234 observations in
`ggplot2::mpg`, with its axis text and legend kept at their intended sizes:

```{r geometry-example}
geometry_colours <- setNames(
  figspec_palette("okabe_ito")[c(6, 7, 4)],
  c("4", "f", "r")
)
geometry_shapes <- setNames(figspec_shapes(3), c("4", "f", "r"))

geometry_example <- ggplot(
  ggplot2::mpg,
  aes(displ, hwy, colour = drv, shape = drv)
) +
  geom_point(alpha = 0.72, size = 1.8) +
  scale_colour_manual(values = geometry_colours) +
  scale_shape_manual(values = geometry_shapes) +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy",
       colour = "Drive type",
       shape = "Drive type") +
  theme_spec("cell_press", base_size = 8) +
  theme(legend.position = "bottom")

geometry_example <- fig_panel_size(
  geometry_example, width = 62, height = 45, units = "mm"
)
sized_geometry <- fig_geometry(geometry_example)
sized_geometry
```

Read the diagram from the inside out. The white rectangle is the 62 by 45 mm
data panel. The surrounding grey area shows the space required by everything
outside the data:

* The `r round(sized_geometry$left_mm, 1)` mm on the left contains the y-axis
  tick labels and axis title.
* The `r round(sized_geometry$bottom_mm, 1)` mm below the panel contains the
  x-axis text and bottom legend, which is why it is the largest allowance.
* The `r round(sized_geometry$top_mm, 1)` mm above and
  `r round(sized_geometry$right_mm, 1)` mm to the right are mainly outer
  spacing.

```{r geometry-diagram, echo = FALSE, results = "asis"}
embed_plot(
  function() print(plot(sized_geometry)),
  width = 5,
  height = 3.6,
  dpi = 200,
  alt = paste(
    "A diagram showing the measured panel and the space used by labels",
    "and margins on each side."
  ),
  out_width = "75%"
)
```

These measurements turn a vague problem such as “the plot looks cramped” into
a specific choice. Shortening y-axis labels can recover horizontal room;
moving or removing the legend can recover vertical room. If those elements are
needed, figspec shows exactly how much larger the canvas must be to preserve
them.

## Panel-sizing functions at a glance

| What you need | Function |
|---|---|
| Set the overall image size | `fig_save(width = ...)` |
| Set an exact plotting-panel size | `fig_panel_size()` or `fig_save(panel_width = ...)` |
| Find one panel width that fits a complete set | `fig_panel_width()` |
| Inspect the current canvas, panel and surrounding space | `fig_geometry()` |
| Check whether a set is aligned | `submission_check()` |

The sections below demonstrate each part of this workflow with real plots and
exported files.

## Keep long labels from shrinking the data area

Both plots below are 85 mm wide and use the same 234 observations, variables
and scales from `ggplot2::mpg`. The only difference is the wording of the
y-axis tick labels. Because the longer labels keep their full text size, they
take space away from the rectangle in which the data are drawn.

The pale blue rectangle marks that data panel; the grey outline marks the
complete image. Colour and shape both identify the vehicle drive type, so the
groups remain distinguishable without relying on colour alone.

```{r create-panel-examples}
drive_colours <- setNames(
  figspec_palette("okabe_ito")[c(6, 7, 4)],
  c("4", "f", "r")
)
drive_shapes <- setNames(figspec_shapes(3), c("4", "f", "r"))

base_plot <- ggplot(
  ggplot2::mpg,
  aes(displ, hwy, colour = drv, shape = drv)
) +
  geom_point(alpha = 0.72, size = 1.8) +
  scale_colour_manual(values = drive_colours) +
  scale_shape_manual(values = drive_shapes) +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy",
       colour = "Drive type",
       shape = "Drive type") +
  theme_spec("cell_press", base_size = 8) +
  theme(
    legend.position = "bottom",
    panel.background = element_rect(
      fill = "#F3F8FC", colour = "#0072B2", linewidth = 0.7
    ),
    plot.background = element_rect(
      fill = "white", colour = "#9AA4AE", linewidth = 0.6
    )
  )

figs <- list(
  short_labels = base_plot,
  labels_written_out = base_plot +
    scale_y_continuous(labels = function(x) paste(x, "miles per gallon"))
)
```

The more polished appearance comes from figspec's `theme_spec()`,
`figspec_palette()` and `figspec_shapes()` helpers. The panel-sizing functions
do not recolour or redesign a chart; they control its physical geometry. This
guide uses the styling and sizing tools together because a finished figure
should be both clear and correctly sized.

### Before: equal canvases, unequal panels

```{r comparison-helper, include = FALSE}
draw_pair <- function(plots, widths_mm, labels, panel_widths_mm,
                      canvas_height_mm = 62, gap_mm = 8) {
  grobs <- lapply(plots, function(x) {
    if (inherits(x, "ggplot")) ggplotGrob(x) else x
  })
  grid::grid.newpage()
  device_width <- grid::convertWidth(
    grid::unit(1, "npc"), "mm", valueOnly = TRUE
  )
  total_width <- sum(widths_mm) + gap_mm * (length(grobs) - 1)
  left <- max(0, (device_width - total_width) / 2)

  for (i in seq_along(grobs)) {
    x <- left + sum(widths_mm[seq_len(i - 1)]) + gap_mm * (i - 1)
    grid::grid.text(
      paste0(labels[[i]], "\nPanel width: ",
             round(panel_widths_mm[[i]], 1), " mm"),
      x = grid::unit(x + widths_mm[[i]] / 2, "mm"),
      y = grid::unit(canvas_height_mm + 11, "mm"),
      gp = grid::gpar(fontsize = 10, fontface = "bold")
    )
    grid::pushViewport(grid::viewport(
      x = grid::unit(x, "mm"), y = grid::unit(2, "mm"),
      width = grid::unit(widths_mm[[i]], "mm"),
      height = grid::unit(canvas_height_mm, "mm"),
      just = c("left", "bottom")
    ))
    grid::grid.draw(grobs[[i]])
    grid::popViewport()
  }
}

draw_stack <- function(plots, widths_mm, labels, panel_widths_mm,
                       canvas_height_mm = 56, gap_mm = 8) {
  grobs <- lapply(plots, function(x) {
    if (inherits(x, "ggplot")) ggplotGrob(x) else x
  })
  grid::grid.newpage()
  device_width <- grid::convertWidth(
    grid::unit(1, "npc"), "mm", valueOnly = TRUE
  )
  device_height <- grid::convertHeight(
    grid::unit(1, "npc"), "mm", valueOnly = TRUE
  )
  row_height <- canvas_height_mm + 10
  total_height <- length(grobs) * row_height +
    (length(grobs) - 1) * gap_mm
  bottom <- max(0, (device_height - total_height) / 2)

  for (i in seq_along(grobs)) {
    y <- bottom + (length(grobs) - i) * (row_height + gap_mm)
    x <- max(0, (device_width - widths_mm[[i]]) / 2)
    grid::grid.text(
      paste0(
        labels[[i]], "\nCanvas: ", round(widths_mm[[i]], 1),
        " mm  |  each panel: ", round(panel_widths_mm[[i]], 1), " mm"
      ),
      x = grid::unit(device_width / 2, "mm"),
      y = grid::unit(y + canvas_height_mm + 6, "mm"),
      gp = grid::gpar(fontsize = 10, fontface = "bold")
    )
    grid::pushViewport(grid::viewport(
      x = grid::unit(x, "mm"), y = grid::unit(y, "mm"),
      width = grid::unit(widths_mm[[i]], "mm"),
      height = grid::unit(canvas_height_mm, "mm"),
      just = c("left", "bottom")
    ))
    grid::grid.draw(grobs[[i]])
    grid::popViewport()
  }
}

embed_png_pair <- function(paths, widths, alt) {
  images <- vapply(seq_along(paths), function(i) {
    sprintf(
      paste0(
        '<img src="data:image/png;base64,%s" ',
        'style="width:%s;max-width:100%%;height:auto" alt="%s">'
      ),
      sub("^data:image/png;base64,", "", knitr::image_uri(paths[[i]])),
      widths[[i]], alt[[i]]
    )
  }, character(1))
  knitr::asis_output(paste0(
    '<div class="figspec-export-pair">',
    paste(images, collapse = "\n"),
    '</div>'
  ))
}

```

```{r panel-before, echo = FALSE, results = "asis"}
before_widths <- fig_panel_width(
  figs,
  spec = "cell_press",
  column = "single",
  format = "tiff"
)

embed_plot(
  function() draw_pair(
    figs,
    widths_mm = c(85, 85),
    labels = c("Numeric tick labels", "Long tick labels"),
    panel_widths_mm = attr(before_widths, "per_figure")
  ),
  width = 7.15,
  height = 3.05,
  dpi = 200,
  alt = paste(
    "Two scatter plots with equal 85 millimetre canvases.",
    "The plot with long y-axis labels has a visibly narrower blue data",
    "panel than the plot with numeric labels."
  )
)
```

The outer canvases are the same width, but the blue panel on the right is
narrower. Its longer labels occupy
`r round(diff(range(attr(before_widths, "per_figure"))), 1)` mm more horizontal
space, leaving less room for the data.

## After: make both data panels the same width

To align a set, first find the widest panel that every figure can share within
the available canvas. `fig_panel_width()` measures the space used by each
figure's labels, axes, legend and margins, then returns the common width that
fits them all:

```{r}
shared_width <- fig_panel_width(
  figs,
  spec = "cell_press",
  column = "single",
  format = "tiff"
)
shared_width
```

The per-figure values show how wide each panel could have been on its own. The
smaller value becomes the shared width because it is the only value that fits
both figures.

Apply that width to every plot with `fig_panel_size()`:

```{r}
aligned <- lapply(figs, fig_panel_size, width = shared_width)
```

```{r panel-after, echo = FALSE, results = "asis"}
after_geometry <- lapply(aligned, fig_geometry)

embed_plot(
  function() draw_pair(
    aligned,
    widths_mm = vapply(
      after_geometry, function(x) x$canvas_width_mm, numeric(1)
    ),
    labels = c("Numeric tick labels", "Long tick labels"),
    panel_widths_mm = vapply(
      after_geometry, function(x) x$panel_width_mm, numeric(1)
    )
  ),
  width = 7.15,
  height = 3.05,
  dpi = 200,
  alt = paste(
    "The same two scatter plots after alignment. Both blue data panels",
    "have the same measured width. The canvas around the long-label plot",
    "is wider so its labels fit without shrinking the data panel."
  )
)
```

The matching blue rectangles show the result: both data panels are now
`r round(shared_width, 1)` mm wide. figspec gives the second figure a wider
canvas, preserving both the shared data area and the full labels.

```{r panel-measurements, echo = FALSE}
comparison <- data.frame(
  Figure = c("Numeric tick labels", "Long tick labels"),
  `Panel before (mm)` = round(attr(before_widths, "per_figure"), 1),
  `Panel after (mm)` = round(vapply(
    after_geometry, function(x) x$panel_width_mm, numeric(1)
  ), 1),
  `Canvas after (mm)` = round(vapply(
    after_geometry, function(x) x$canvas_width_mm, numeric(1)
  ), 1),
  check.names = FALSE
)
knitr::kable(comparison, row.names = FALSE, align = c("l", "r", "r", "r"))
```

The table is the reproducible check behind the illustration. Before alignment,
the panels differ by
`r round(diff(range(attr(before_widths, "per_figure"))), 1)` mm. After
alignment, both measure `r round(shared_width, 1)` mm; the outer canvases are
allowed to differ so the labels retain their intended size.

`submission_check()` measures the same alignment across a complete set:

```{r}
submission_check(aligned, "cell_press")
```

Publisher profiles specify the overall image dimensions, not the size of the
plotting area inside the image. `submission_check()` therefore presents panel
alignment as layout information alongside the publication checks. The report
also identifies any file properties that should be checked after export.

For a report, thesis, presentation or dashboard, supply the available canvas
width directly:

```{r}
report_panel_width <- fig_panel_width(figs, width = 150, units = "mm")
report_panel_width
```

## Set an exact data area

Sometimes the data region itself must be a known size: for example, maps that
will be compared at the same scale, plots that must align in a report, or a
series whose axes should occupy identical space. `fig_panel_size()` sets that
region directly:

```{r}
sized <- fig_panel_size(
  figs$short_labels,
  width = 62,
  height = 45,
  units = "mm"
)
fig_geometry(sized)
```

The data panel is exactly 62 by 45 mm. figspec measures the labels, legend and
margins, then gives the complete figure the additional space they require.
Nothing in the data or scales is stretched to make the dimensions fit.

## Export files with the same data area

`fig_save()` applies the panel dimensions while it writes the file. This avoids
the usual cycle of exporting, measuring, changing the canvas and exporting
again. Here, two real PNG files are written at 300 dpi. Their labels differ,
but both receive a 62 by 45 mm data panel:

```{r}
short_file <- tempfile(fileext = ".png")
long_file <- tempfile(fileext = ".png")

short_saved <- fig_save(
  short_file,
  figs$short_labels + labs(title = "Short tick labels"),
  panel_width = 62,
  panel_height = 45,
  dpi = 300
)
long_saved <- fig_save(
  long_file,
  figs$labels_written_out + labs(title = "Long tick labels"),
  panel_width = 62,
  panel_height = 45,
  dpi = 300
)
```

```{r exported-files, echo = FALSE, results = "asis"}
embed_png_pair(
  c(short_saved, long_saved),
  c("43%", "52%"),
  c(
    "The exported short-label PNG with a 62 by 45 millimetre blue data panel.",
    paste(
      "The exported long-label PNG with the same 62 by 45 millimetre blue",
      "data panel and a wider surrounding canvas."
    )
  )
)
```

```{r exported-file-check, echo = FALSE}
export_paths <- list(short_saved, long_saved)
export_geometry <- lapply(export_paths, fig_geometry)
export_reports <- lapply(export_paths, fig_check, spec = "cell_press")
actual_value <- function(report, item) {
  trimws(report$actual[report$check == item])
}

export_summary <- data.frame(
  File = c("Short labels", "Long labels"),
  `Panel (mm)` = vapply(
    export_geometry,
    function(x) paste0(round(x$panel_width_mm, 1), " x ",
                       round(x$panel_height_mm, 1)),
    character(1)
  ),
  `Canvas (mm)` = vapply(
    export_reports,
    function(x) paste(actual_value(x, "Width"), "x",
                       actual_value(x, "Height")),
    character(1)
  ),
  Resolution = vapply(
    export_reports, actual_value, character(1), item = "Resolution"
  ),
  check.names = FALSE
)
knitr::kable(export_summary, row.names = FALSE, align = c("l", "c", "c", "c"))
```

These values come from the files that were actually written: `fig_check()`
reopened each PNG and read its saved dimensions and resolution. The panels
match, and the long-label image receives the additional canvas needed to keep
its text at the intended size.

For a publication profile, add `spec` and `column`. figspec will retain the
publisher's required canvas width while fitting the requested panel inside it:

```{r, eval = FALSE}
fig_save(
  "figure_1.tiff",
  figs$short_labels,
  spec = "cell_press",
  column = "single",
  panel_width = shared_width
)
```

## Use all the available canvas

If a figure must fit an 85 mm space, `panel_width = "max"` measures the labels,
axes, legend and margins, then gives all remaining room to the data panel. The
comparison below shows how much space this recovers from the same 85 mm canvas:

```{r}
conservative_file <- tempfile(fileext = ".png")
maximum_file <- tempfile(fileext = ".png")

conservative_panel <- fig_save(
  conservative_file,
  figs$short_labels + labs(title = "A guessed 55 mm panel"),
  width = 85,
  panel_width = 55,
  panel_height = 42,
  dpi = 300
)
maximum_panel <- fig_save(
  maximum_file,
  figs$short_labels + labs(title = "The widest panel that fits"),
  width = 85,
  panel_width = "max",
  panel_height = 42,
  dpi = 300
)
```

```{r maximum-panel-comparison, echo = FALSE, results = "asis"}
embed_png_pair(
  c(conservative_panel, maximum_panel),
  c("48%", "48%"),
  c(
    "An 85 millimetre figure with a conservatively chosen 55 millimetre data panel.",
    paste(
      "The same 85 millimetre figure using the widest data panel that its",
      "labels and legend allow."
    )
  )
)
```

```{r maximum-panel-table, echo = FALSE}
maximum_comparison <- data.frame(
  Method = c("Guessed panel width", 'panel_width = "max"'),
  `Canvas width (mm)` = round(vapply(
    list(conservative_panel, maximum_panel),
    function(x) fig_geometry(x)$canvas_width_mm,
    numeric(1)
  ), 1),
  `Panel width (mm)` = round(vapply(
    list(conservative_panel, maximum_panel),
    function(x) fig_geometry(x)$panel_width_mm,
    numeric(1)
  ), 1),
  check.names = FALSE
)
knitr::kable(maximum_comparison, row.names = FALSE, align = c("l", "r", "r"))
```

`"max"` is calculated for the particular figure, so it makes full use of the
canvas without clipping its surrounding material. It is ideal for one figure.
To give several different figures a common panel width, continue to use
`fig_panel_width()` as shown earlier.

## Setting both canvas and panel dimensions

Canvas width and panel width describe different parts of the same figure:

```
canvas = panel + axes + labels + legend + margins
```

When you provide both, figspec measures the surrounding material and checks
that the requested combination is possible:

```{r}
both_sizes <- fig_save(
  tempfile(fileext = ".png"),
  figs$short_labels,
  width = 120,
  panel_width = 62,
  dpi = 300
)
fig_geometry(both_sizes)
```

When the requested dimensions need more room than the canvas provides, figspec
stops before writing the file. The message gives the canvas width or panel
width that will make the combination work:

```{r, error = TRUE}
fig_save(
  tempfile(fileext = ".png"),
  figs$short_labels,
  width = 65,
  panel_width = 62,
  dpi = 300
)
```

## Keep every panel readable in a multi-panel figure

### Faceted plots

Facets divide one canvas among several data panels. In this example, an 85 mm
canvas leaves about 24 mm for each of three panels. Setting each panel to 30 mm
makes figspec calculate the wider canvas needed for the complete figure.

```{r faceted-plot}
facet_colours <- setNames(
  figspec_palette("okabe_ito")[c(6, 7, 4)],
  c("4", "6", "8")
)

faceted <- ggplot(subset(ggplot2::mpg, cyl %in% c(4, 6, 8)),
                  aes(displ, hwy, colour = factor(cyl))) +
  geom_point(alpha = 0.72, size = 1.7) +
  scale_colour_manual(values = facet_colours) +
  guides(colour = "none") +
  facet_wrap(~ cyl, nrow = 1) +
  labs(x = "Engine displacement (litres)",
       y = "Highway fuel economy") +
  theme_spec("cell_press", base_size = 8) +
  theme(
    panel.background = element_rect(
      fill = "#F3F8FC", colour = "#0072B2", linewidth = 0.7
    ),
    plot.background = element_rect(
      fill = "white", colour = "#9AA4AE", linewidth = 0.6
    )
  )
```

```{r}
faceted_width_before <- fig_panel_width(
  list(faceted = faceted), width = 85, units = "mm"
)
faceted_before <- fig_panel_size(
  faceted, width = faceted_width_before, height = 42
)
faceted_sized <- fig_panel_size(faceted, width = 30, height = 42)
```

```{r faceted-before-after, echo = FALSE, results = "asis"}
faceted_geometry <- lapply(
  list(before = faceted_before, after = faceted_sized),
  fig_geometry
)
embed_plot(
  function() draw_stack(
    list(faceted_before, faceted_sized),
    widths_mm = vapply(
      faceted_geometry, function(x) x$canvas_width_mm, numeric(1)
    ),
    labels = c("Before: three panels share 85 mm",
               "After: every panel receives 30 mm"),
    panel_widths_mm = vapply(
      faceted_geometry, function(x) x$panel_width_mm, numeric(1)
    )
  ),
  width = 5.8,
  height = 5.8,
  dpi = 200,
  alt = paste(
    "A before-and-after comparison of a three-facet plot. On the 85",
    "millimetre canvas each panel is 24.1 millimetres wide; after sizing,",
    "every panel is 30 millimetres wide and the canvas expands to fit."
  )
)
```

With conventional canvas sizing, R honours the requested 85 mm image width and
leaves `r round(faceted_geometry[[1]]$panel_width_mm, 1)` mm for each facet.
figspec lets you set every facet to 30 mm and calculates that the complete
image must be
`r round(faceted_geometry[[2]]$canvas_width_mm, 1)` mm wide.

The data ranges and text sizes stay the same. Every facet now has a known
physical width, making the complete chart easier to plan and read.

### Separate plots assembled with patchwork

figspec can measure through a patchwork composition and control the physical
size of every underlying data panel in the finished file. This makes an
assembled figure subject to the same reproducible sizing rules as a single
plot.

Suppose this composition is first exported on a 120 mm canvas. Its two panels
are about 45 mm wide. If a report layout requires each panel to be 60 mm wide,
figspec can set that measurement directly and calculate the canvas required
for the axes, titles, spacing and margins.

```{r patchwork-composition}
vehicle_classes <- sort(unique(ggplot2::mpg$class))
class_colours <- setNames(
  figspec_palette("viridis", length(vehicle_classes)),
  vehicle_classes
)

scatter_panel <- figs$short_labels +
  guides(colour = "none", shape = "none") +
  labs(title = "Fuel economy and engine size")

box_panel <- ggplot(ggplot2::mpg, aes(hwy, class, fill = class)) +
  geom_boxplot(show.legend = FALSE, width = 0.65, outlier.alpha = 0.45) +
  scale_fill_manual(values = class_colours) +
  labs(title = "Fuel economy by vehicle class",
       x = "Highway fuel economy", y = NULL) +
  theme_spec("cell_press", base_size = 8) +
  theme(
    panel.background = element_rect(
      fill = "#F8F6F1", colour = "#D55E00", linewidth = 0.7
    ),
    plot.background = element_rect(
      fill = "white", colour = "#9AA4AE", linewidth = 0.6
    )
  )

composition <- patchwork::wrap_plots(scatter_panel, box_panel, nrow = 1)
composition_width_before <- fig_panel_width(
  list(composition = composition), width = 120, units = "mm"
)
composition_before <- fig_panel_size(
  composition, width = composition_width_before, height = 42
)
composition_after <- fig_panel_size(composition, width = 60, height = 42)
```

```{r patchwork-before-after, echo = FALSE, results = "asis"}
composition_geometry <- lapply(
  list(composition_before, composition_after), fig_geometry
)
embed_plot(
  function() draw_stack(
    list(composition_before, composition_after),
    widths_mm = vapply(
      composition_geometry, function(x) x$canvas_width_mm, numeric(1)
    ),
    labels = c("Before: two panels share 120 mm",
               "With figspec: each panel is exactly 60 mm"),
    panel_widths_mm = vapply(
      composition_geometry, function(x) x$panel_width_mm, numeric(1)
    )
  ),
  width = 7.2,
  height = 5.8,
  dpi = 200,
  alt = paste(
    "A comparison of a two-plot composition at two physical sizes.",
    "A 120 millimetre canvas produces panels about 45 millimetres wide;",
    "figspec gives each panel 60 millimetres and calculates the larger",
    "canvas required."
  )
)
```

The compositions retain the same data, scales and relative arrangement, so
the comparison isolates the change in geometry. Each data panel grows from
`r round(composition_geometry[[1]]$panel_width_mm, 1)` to
`r round(composition_geometry[[2]]$panel_width_mm, 1)` mm, and figspec expands
the canvas from `r round(composition_geometry[[1]]$canvas_width_mm, 1)` to
`r round(composition_geometry[[2]]$canvas_width_mm, 1)` mm without shrinking
the labels.

The result can then be passed directly to `fig_save()` for export or compared
with other figures through `submission_check()`. The composition now has
known, repeatable panel dimensions that can be reproduced in every export.


```{r, include = FALSE}
unlink(c(short_file, long_file, conservative_file, maximum_file))
```
