---
title: "Choose, Verify and Add Specifications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choose, Verify and Add Specifications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(figspec)
```

## Why figspec keeps a specification registry

Requirements for figures may come from a publisher, an individual journal, an
organisation or a project. These requirements might change from time to time. As such they may need to be updated. Additionally, the number of journals grows and more need to be added.

figspec keeps these requirements in a registry to make it easy to reuse a specification without needing to look up changes to every specific requirement everytime you build figures in R. The registry also keeps the source and review date with the specification, making it possible to see
what figspec is applying and to update the record when the guidance changes.

## Choose the right specification

Start with the specification that most closely matches the work you are
preparing. If the registry contains an entry for the individual journal, use
that entry. If it does not, a publisher-wide entry may be appropriate when the
journal directs authors to the publisher's general guidance.

Publisher-wide guidance does not necessarily cover every requirement for every
journal in the publisher's portfolio. An individual journal may use different
page dimensions, accept different file formats or add instructions of its own.
Check the journal's guidance before relying on a publisher-wide profile.

`spec_list()` lists the specifications currently available. This example shows
one publisher-wide profile and one journal-specific profile:

```{r choose-specification}
available <- spec_list()
available[
  match(c("elsevier", "jid"), available$id),
  c("id", "name", "publisher", "origin")
]
```

If no suitable profile exists, you can add a specification for a publication,
organisation or project. figspec does not automatically join a publisher-wide
profile to separate journal instructions. When both apply, add a specification
that records the combined requirements and the sources from which they came.

## Check where a specification came from

Every specification included with figspec records the page from which its
requirements were taken and the date that page was last checked. You can see both
with `spec_get()`:

```{r inspect-source}
spec <- spec_get("elsevier")
spec$source_url
spec$verified_on
```

The source link lets you read the original guidance and confirm that it applies
to the work you are preparing. The review date tells you when the registry
entry was last compared with that guidance. It does not mean that the source
has remained unchanged since that date or that every statement has been
interpreted correctly.

When the guidance gives a numerical requirement, the specification can also
keep the passage that supports the recorded value. For example, the widths in
the Elsevier profile are accompanied by the wording from its guidance:

```{r inspect-source-quote}
writeLines(strwrap(spec$source_quote_width))
```

Some publications place their requirements on several pages. A specification
can therefore record more than one source and show which page supports each
requirement. This is useful when dimensions, resolution and text requirements
are documented separately.

## Understand what is included and what is missing

Publications do not always state a requirement for every aspect of a figure. When a
value is missing from a specification, the source may have been checked without
finding that requirement, or that part of the source may not have been reviewed
yet. The difference matters: a requirement that was not found is not the same
as a requirement that nobody has looked for.

figspec keeps these situations separate and reports them as follows:

| What was found | What it means | How figspec reports it |
|---|---|---|
| The reviewed guidance states a requirement | The requirement is recorded and can be checked when the figure or file provides enough information | `pass`, `fail` or `unknown` |
| The relevant guidance was checked but does not state a requirement | No requirement was found in the sources reviewed for that specification | Not specified by the source |
| The requirement has not yet been reviewed | figspec does not yet know whether the source states it | Not yet harvested |

`registry_status()` shows how many requirements fall into each group for every
profile. This example compares two profiles in the registry:

```{r registry-field-status}
status <- registry_status()
status[
  status$id %in% c("cell_press", "aps"),
  c("id", "stated", "confirmed_absent", "unharvested")
]
```

These counts describe how much of each profile has been reviewed. They do not
describe how many requirements a publisher ought to have. *Confirmed absent*
means only that the sources checked for that specification did not state the
requirement. Another page or a later version of the guidance may still contain
it.

## Add your own specification

If the registry does not include the specification you need, you can add it.
The method depends on whether you need the specification only for your current
work, want to reuse it in other projects, or want to propose it for inclusion
with figspec.

| What you want to do | Function | How long the specification remains available |
|---|---|---|
| Use a specification in the current R session | `spec_register()` | Until the R session ends |
| Keep a specification for a project or organisation | `spec_load()` | Whenever its YAML file is loaded |
| Propose a specification for a future figspec release | `registry_entry_template()` | After it has been reviewed and accepted into the package |

The function names still use *journal* because they were introduced when the
registry focused on journal requirements. The specifications themselves are
not limited to journals. They can describe a publisher, report, thesis,
presentation or another type of project.

### Add a specification for the current session

Use `spec_register()` to define a specification directly in R. This example
adds the requirements for an internal report and then retrieves its 170 mm
double-width value:

```{r register-session-specification}
spec_register(
  id = "lab_report",
  name = "Our lab report format",
  source_url = "internal:handbook-v3",
  verified_on = "2026-09-04",
  requirements = list(
    columns = list(single = 100, double = 170),
    font_min_pt = 9,
    formats = c("png", "pdf")
  )
)

fig_width("lab_report", "double")
spec_list()[spec_list()$id == "lab_report", c("id", "name", "origin")]
```

The result is marked with the origin `user` because it was added during the R
session rather than supplied with figspec. This label identifies where the
entry came from; it is not a rating of its quality.

### Save a specification for reuse

To reuse a specification, keep it in a YAML registry file that can be stored
with the project or shared by an organisation. `registry_entry_template()` prints a
template containing the information that figspec can record. This example runs
the function and shows the beginning of that template:

```{r new-entry-template}
template <- capture.output(registry_entry_template(
  "plos_biology",
  "PLOS Biology",
  "https://journals.plos.org/plosbiology/s/figures"
))
writeLines(c(template[1:12], "  ..."))
```

Add a requirement only when it is supported by the source. List it under
`not_stated:` only when the relevant guidance has been checked and does not
state that requirement. Leave it out of both places when it has not yet been
reviewed.

The following example writes a real temporary registry file, checks it, loads
it and retrieves the width recorded in it:

```{r validate-and-load-registry}
registry_file <- tempfile(fileext = ".yaml")
writeLines(c(
  "schema_version: 2",
  "journals:",
  "- id: team_report",
  "  name: Research team report",
  "  publisher: Internal",
  "  disciplines: [research]",
  "  source_url: internal:report-handbook-v4",
  "  verified_on: '2026-09-04'",
  "  requirements:",
  "    columns: {full: 160}",
  "    source_quote_width: Figures occupy the 160 mm report content width.",
  "    dpi_min: 300",
  "    source_quote_dpi: Export raster figures at 300 dpi or higher.",
  "    formats: [png, pdf]",
  "    font_min_pt: 9",
  "    source_quote_font: Figure text must be at least 9 pt.",
  "  not_stated: [height_max_mm]"
), registry_file)

registry_validate_file(registry_file)
loaded_ids <- spec_load(registry_file)
loaded_ids
fig_width("team_report", "full")
spec_list()[
  spec_list()$id == "team_report",
  c("id", "name", "dpi_min", "origin")
]
```

```{r cleanup-registry-file, include = FALSE}
unlink(registry_file)
```

Validation checks that the file can be read and that its entries use the
expected structure and values. It cannot confirm that the source was
interpreted correctly. That still needs to be reviewed by a person.

### Propose a specification for figspec

To propose a specification for inclusion with figspec, begin with the same
template, include the source wording that supports each requirement and check
the completed file with `registry_validate_file()`. A maintainer can then
review the source and the values before deciding whether to include the profile
in a future release.

## Keep specifications current

Requirements can change after a specification has been added to the registry.
The `verified_on` date shows when its source was last checked. Use
`registry_stale_entries()` to find specifications that have not been checked within a
period you choose:

```{r stale-profiles}
registry_stale_entries(max_age_days = 365)
```

A specification listed by this function is not necessarily wrong. It is due
for another comparison with its source. If the guidance has changed, update
the recorded requirements, the passages that support them and the review date
together.

`registry_check_sources()` checks whether the web links recorded in the registry still
respond. It is not run while this page is built because it requires an internet
connection and some publisher websites block automated requests. A working
link shows only that the page can be reached. Someone still needs to read the
page and check whether its guidance has changed.

## Contribute to the bundled registry

The bundled registry is included with figspec and is available to everyone who
installs the package. A new or revised profile may therefore affect many users.
Each contribution needs to identify the source, record when it was last checked
and include the wording that supports important requirements.

### Decide whether the guidance states a requirement

Publication guidance often combines instructions, preferences and general
advice. Not every useful statement should become a requirement that figspec
uses to pass or fail a figure.

| Example wording | How it is recorded | Why |
|---|---|---|
| PLOS: “Use only Arial, Times, or Symbol font” | Requirement | The instruction allows only the named fonts |
| OUP: “at least 300 dpi” within a guide described as tips | Requirement | The statement gives a measurable minimum |
| Elsevier: “no smaller than 6 pt” | Requirement | The statement gives a minimum text size |
| Sage: “We recommend having no more than 7 series” | Recommendation | The advice is useful but is not stated as mandatory |
| ACS: “Helvetica or Arial fonts work well” | Not recorded as a font restriction | The statement does not say that other fonts are prohibited |

Read the complete instruction and its surrounding explanation before deciding
how to record it. If the wording does not clearly require the author to meet a
condition, do not turn it into a checkable requirement.

Keep personal or organisational design preferences in a house style rather
than recording them as publication requirements. A preferred grid, background
or spacing choice is not a requirement unless the cited source states that it
is.

### Validate and review the entry

Run `registry_validate_file()` before submitting a new or revised registry
file. It checks that the file contains the required information, that values
use the expected formats and that the same item has not been recorded as both
stated and not stated.

Passing validation does not prove that the correct source was used or that its
wording was interpreted correctly. A maintainer must still compare the entry
with the cited guidance and review the supporting passages.

### Use harvesting tools as a starting point

The GitHub repository contains a
[`data-raw/harvest.R`](https://github.com/dansemakula/figspec/blob/HEAD/data-raw/harvest.R)
toolkit that can help maintainers find publication pages, retrieve current or
archived guidance and identify statements that may contain requirements. It is
a development tool and is not installed with the CRAN package.

The toolkit does not add requirements to the registry automatically. A person
must confirm that the page applies to the intended publication, read the
guidance in context, keep the wording that supports each value and decide what
is stated, not stated or still unreviewed. The completed entry can then be
validated and submitted for review by following the
[contribution guide](https://github.com/dansemakula/figspec/blob/HEAD/CONTRIBUTING.md).
