---
title: "Build and check tables from R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build and check tables from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
library(figspec)
library(ggplot2)
```

Tables often have requirements of their own. A report may require 9 point
Arial text and no vertical lines. A journal may require an editable Word table,
a portrait page, a short title and abbreviations explained below the table.
These are not figure rules, and figspec keeps them separate.

This guide shows the complete table workflow: describe the table you need,
apply the requirements while it is editable, export it, and check the file
that was actually written.

## Describe the table you need

A table specification uses a `tables` section. This example defines the
requirements for a research report directly in R, so you can see exactly what
will be applied:

```{r table-spec}
report_spec <- list(
  name = "Clinical research report",
  source_url = "internal:report-table-guide",
  verified_on = as.character(Sys.Date()),
  tables = list(
    formats = c("html", "docx"),
    orientation = "portrait",
    font_families = "Arial",
    font_min_pt = 9,
    header_bold = TRUE,
    vertical_rules = FALSE,
    horizontal_rules = "minimal",
    repeat_header = TRUE
  )
)

table_spec(report_spec)
```

The same specification can be kept with the project and reused in later R
sessions:

```{r save-table-specification}
table_spec_file <- file.path(tempdir(), "clinical-report-specifications.yml")
spec_save(report_spec, table_spec_file, id = "clinical_research_report")
spec_load(table_spec_file)
```

The saved entry retains both table and figure requirements. Loading it makes
the project specification available in the current R session while leaving
the installed package unchanged.

Bundled publication profiles can be used in the same way when their table
requirements have been reviewed. `registry_status()` shows table coverage
separately from figure coverage; an empty table section means that figspec has
not yet reviewed that profile's table guidance.

## Apply the requirements before export

We will use 12 real observations from `ggplot2::mpg`. The plain data frame is
useful data, but it does not yet record a font size, header treatment or rule
style.

```{r plain-table}
vehicle_table <- mpg[1:12, c(
  "manufacturer", "model", "displ", "year", "hwy"
)]

knitr::kable(vehicle_table)
```

`table_apply_spec()` turns the data frame into an editable gt table and applies
the requirements the table system can express:

```{r styled-table}
styled_table <- table_apply_spec(vehicle_table, report_spec)
styled_table
```

The rows and values have not changed. The difference is that the editable
table now carries the required type size, font, header emphasis and rule
choices. `table_check()` can inspect those properties before any file is
written:

```{r check-live-table}
table_check(styled_table, report_spec)
```

At this stage, the live-table report covers the styling and structure that can
already be inspected. The file format is checked after export, when a completed
file is available.

## Export and check the completed file

`table_save()` exports through the table's own system. It first writes a
temporary file in the destination directory, reopens and checks it, and only
then places it at the requested path.

```{r save-table}
html_path <- file.path(tempdir(), "vehicle-summary.html")
saved_table <- table_save(
  html_path,
  styled_table,
  spec = report_spec,
  transform = FALSE
)

attr(saved_table, "figspec_table_report")
```

The combined report now contains both kinds of evidence: styling from the
editable table and HTML format and validity from the completed file. Use DOCX
instead when an editable Word file is required:

```{r save-docx, eval=FALSE}
word_path <- file.path(tempdir(), "vehicle-summary.docx")
table_save(word_path, vehicle_table, report_spec)
```

## Use the table system you already work with

Continue using the table system that suits your work. figspec accepts:

* data frames and matrices, which become gt tables;
* existing gt tables;
* flextable objects, including Word-oriented workflows;
* knitr and kableExtra tables; and
* grid tables, which can be exported as PDF or images.

The same public calls are used for all of them:

```{r backend-shape, eval=FALSE}
table_apply_spec(my_table, report_spec)
table_save("my-table.docx", my_table, report_spec)
table_check("my-table.docx", report_spec)
```

The output formats still depend on the table system:

| Table supplied to figspec | Available output |
|---|---|
| data frame or matrix | HTML, DOCX, RTF, TeX, PDF or PNG |
| gt | HTML, DOCX, RTF, TeX, PDF or PNG |
| flextable | HTML, DOCX, RTF or PNG |
| HTML or LaTeX kable | its native format, PDF, PNG or JPEG |
| grid table | PDF, PNG or JPEG |

HTML-to-PDF or image conversion uses a local browser, while LaTeX-to-PDF or
image conversion uses a local LaTeX installation. figspec checks that the
required renderer and output combination are available before writing the
file, so the extension always matches its contents.

## Review figures and tables together

A manuscript usually contains both. Put them in one named list and let
`submission_check()` select the appropriate checks:

```{r mixed-submission}
figure <- ggplot(mpg, aes(displ, hwy, colour = drv, shape = drv)) +
  geom_point() +
  labs(
    x = "Engine displacement (litres)",
    y = "Highway fuel economy",
    colour = "Drive layout",
    shape = "Drive layout"
  )

mixed_spec <- report_spec
mixed_spec$columns <- list(single = 85)
mixed_spec$dpi_min <- 300
mixed_spec$formats <- "png"

mixed_review <- submission_check(
  list(
    fuel_economy_figure = figure,
    vehicle_table = styled_table
  ),
  spec = mixed_spec,
  column = "single"
)

mixed_review[, c("file", "asset", "result")]
```

The summary labels each item as a figure or table. `submission_detail()` opens
the full report for either one.

## Complete the editorial review

Some requirements are measurements: file format, page orientation, width,
font size, header emphasis or the presence of vertical rules. figspec can
apply or inspect those when the table system exposes reliable evidence.

Other requirements ask whether a title is sufficiently short, every
abbreviation is defined or a footnote is clear. Read those parts of the table
in the context of the surrounding report or manuscript. figspec keeps each one
in the report so the editorial review can be completed before delivery.
